---
title: "Gresham Lab Flow Cytometry Multi Timepoint Analysis"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---


## R Markdown

This is the code to generate figures about CNV dynamics from aggregated flow cyto data.


This notebook contains the code necessary to to generate figures from flow cytometry data across timepoints in the Gresham Lab. 

To analyze flow cytometry data, you MUST use the latest version of this code, available on the [Gresham Lab github](https://github.com/GreshamLab/flow).

**Experimental overview**

Write a detailed description of your experiment here including the goal of the analysis and your interpretation of the results.   
If you still see this text it means that you have not described the experiment and whatever follows is meaningless.

**Requirements**

In order to run this code you need:

* have run the "Gresham Lab Floww Cytometry Single Timepoint Analysis" and saved the proportional data, the filtered individual flow data, and/or the statistical data on one or more flowcyto runs


**Output**  
This script combines the data from multiple flow cyto runs and can save the combined runs in the following formats

1. As a dataframe converted from fcs with all or some of the data.
2. As a .csv with summary statistics (e.g. median FL1 per sample)
3. As a .csv with proportional data  (e.g. cells flourescent in population)

**Libraries**

```{r Load libs}

#Load libraries
library(ggcyto)
library(tidyverse)
library(ggjoy)
library(ggforce)

```


```{r User Defined Variables}
#working directory
dir = '.'

#file location
path.data = "/Users/Brandt/Google Drive/MiniStatRun_10_2018/2019-01-24/"
#path.data = "/Users/nathanbrandt/Google Drive/MiniStatRun_10_2018/"

list.folders <- c("LTEE_mCitrine_GAP1_Variants_T00", 
                  "LTEE_mCitrine_GAP1_Variants_T06", 
                  "LTEE_mCitrine_GAP1_Variants_T07", 
                  "LTEE_mCitrine_GAP1_Variants_T08.3", 
                  #"LTEE_mCitrine_GAP1_Variants_T11.1", 
                  "LTEE_mCitrine_GAP1_Variants_T11.2", 
                  #"LTEE_mCitrine_GAP1_Variants_T13.1", 
                  "LTEE_mCitrine_GAP1_Variants_T13.2", 
                  "LTEE_mCitrine_GAP1_Variants_T14", 
                  "LTEE_mCitrine_GAP1_Variants_T15", 
                  "LTEE_mCitrine_GAP1_Variants_T18", 
                  "LTEE_mCitrine_GAP1_Variants_T22", 
                  "LTEE_mCitrine_GAP1_Variants_T25", 
                  "LTEE_mCitrine_GAP1_Variants_T27", 
                  "LTEE_mCitrine_GAP1_Variants_T29", 
                  "LTEE_mCitrine_GAP1_Variants_T34")

#name of the experiment
name <- "LTEE_mCitrine_GAP1_Variants"

#date of extracted data to use
date <- "2019-01-24"


#style of input
#Read in and combine individual Proportional Data
read.prop <- "Yes"
#Read in and combine individual DataFrames 
read.df <- "Yes"
#Read in and combine individual Experiment Statistics
read.stats <- "Yes"

#Save combined Proportional Data
save.prop <- "Yes"
#Save combined DataFrame - Individual data points
save.df <- "Yes"
#Load combined Experiment Statistics
save.stats <- "Yes"

#Load combined Proportional Data
load.prop <- "No"
#Load combined DataFrame - Individual data points
load.df <- "No"
#Load combined Experiment Statistics
load.stats <- "No"


#filenames
file.prop <- paste(name,"_prop_alldata_",Sys.Date(),sep="")
file.df <- paste(name,"_df_alldata_",Sys.Date(),sep="")
file.stats <- paste(name,"_stats_alldata_",Sys.Date(),sep="")

```


```{r Read or load in  Data}

#Read in and Combine Proportional Data
if(read.prop=="Yes"){
  data.prop <- read_csv(paste(path.data,list.folders[1],"_prop_",date,".csv", sep=""))
  for(i in 2:length(list.folders)){
    data.prop_temp <- read_csv(paste(path.data,list.folders[i],"_prop_",date,".csv", sep=""))
    data.prop<- bind_rows(data.prop, data.prop_temp)
  }
  data.prop <- select(data.prop,-c(1))
}
if(save.prop=="Yes"){
  write.csv(data.prop, file= paste(file.prop,".csv",sep=""), row.names=TRUE, quote=F)
}

#Read in and Combine data frames
if(read.df=="Yes"){
  load(paste(path.data,list.folders[1],"_df_",date,".Rdata",sep=""))
  data.df <- filtered.data
  for(i in 2:length(list.folders)) {
    load(paste(path.data,list.folders[i],"_df_",date,".Rdata",sep=""))
    data.df_temp<-filtered.data
    data.df <- rbind(data.df, data.df_temp)
  }
}
if(save.df=="Yes"){
  save(data.df, file=paste(file.df,".Rdata",sep=""))
}

#Read in and Combine Statistical Data
if(read.stats=="Yes"){
  data.stats <- read_csv(paste(path.data,list.folders[1],"_stats_",date,".csv", sep=""))
  for(i in 2:length(list.folders)){
    data.stats_temp <- read_csv(paste(path.data,list.folders[i],"_stats_",date,".csv", sep=""))
    data.stats<- bind_rows(data.stats, data.stats_temp)
  }
  data.stats <- select(data.stats,-c(1))
} 
if(save.stats=="Yes"){
  write.csv(data.stats, file= paste(file.stats,".csv",sep=""), row.names=TRUE, quote=F)
}

#Loading Data
#Read in and Combine Proportional Data
if(load.prop=="Yes"){
  alldata.prop <- read_csv(paste(path.data,name,"_prop_alldata_",date,".csv",sep=""))
}
#Read in and Combine data frames
if(load.df=="Yes"){
  load(paste(path.data,name,"_df_alldata_",date,".Rdata",sep=""))
}
#Read in and Combine Statistical Data
if(load.stats=="Yes"){
  alldata.stats <- read_csv(paste(path.data,name,"_stats_alldata_",date,".csv",sep=""))
}


```

##Plots


```{r Median fl1 over time}


alldata <- data

#median  fluor#
ggplot(alldata, aes(x=TIMEPOINT,y=(NORMALIZED_GFP_MEDIAN), colour=VESSEL)) +
  geom_point() +
  geom_line(aes(linetype=VESSEL), size=1.2) +
  #scale_y_continuous('Normalized fluorescence signal', limits=c(0,0.8)) +
  scale_linetype_manual(values=rep(c("solid", "dashed", "dotdash", "longdash", "twodash", "F1", "4C88C488", "12345678"),4)) +
  #scale_linetype_manual(values=c("twodash", "twodash", "twodash", "solid")) +
  scale_x_continuous("Generations") +
  theme_classic() +
  #scale_color_brewer(palette="Spectral", direction = 1) +
  #scale_color_manual(values = c("#E5F5E0","#C7E9C0", "#A1D99B", "#74C476", "#41AB5D", "#238B45", "#006D2C", "#00441B","#2171B5", "#08519C", "#08306B")) + 
  #facet_grid(VESSEL ~ .) +
  #scale_color_manual(name="Type",values=c("black", "black", "#41ab5d")) +
  facet_wrap_paginate(~VESSEL, ncol = 2, nrow = 2, page = 1) +
  guides(colour = guide_legend(override.aes = list(size=2))) + 
  theme(legend.title = element_text(colour="black", size=15, face="bold"), axis.title.y = element_text(face="bold", size=14), axis.text.y = element_text(size=13), axis.title.x = element_text(face="bold", size=16), axis.text.x = element_text(size=15))

ggplot(alldata, aes(x=TIMEPOINT,y=(NORMALIZED_GFP_MEDIAN), colour=VESSEL)) +
  geom_point() +
  geom_line(aes(linetype=VESSEL), size=1.2) +
  #scale_y_continuous('Normalized fluorescence signal', limits=c(0,0.8)) +
  scale_linetype_manual(values=rep(c("solid", "dashed", "dotdash", "longdash", "twodash", "F1", "4C88C488", "12345678"),4)) +
  #scale_linetype_manual(values=c("twodash", "twodash", "twodash", "solid")) +
  scale_x_continuous("Generations") +
  theme_classic() +
  #scale_color_brewer(palette="Spectral", direction = 1) +
  #scale_color_manual(values = c("#E5F5E0","#C7E9C0", "#A1D99B", "#74C476", "#41AB5D", "#238B45", "#006D2C", "#00441B","#2171B5", "#08519C", "#08306B")) + 
  #facet_grid(VESSEL ~ .) +
  #scale_color_manual(name="Type",values=c("black", "black", "#41ab5d")) +
  facet_wrap_paginate(~VESSEL, ncol = 2, nrow = 2, page = 2) +
  guides(colour = guide_legend(override.aes = list(size=2))) + 
  theme(legend.title = element_text(colour="black", size=15, face="bold"), axis.title.y = element_text(face="bold", size=14), axis.text.y = element_text(size=13), axis.title.x = element_text(face="bold", size=16), axis.text.x = element_text(size=15))

ggplot(alldata, aes(x=TIMEPOINT,y=(NORMALIZED_GFP_MEDIAN), colour=VESSEL)) +
  geom_point() +
  geom_line(aes(linetype=VESSEL), size=1.2) +
  #scale_y_continuous('Normalized fluorescence signal', limits=c(0,0.8)) +
  scale_linetype_manual(values=rep(c("solid", "dashed", "dotdash", "longdash", "twodash", "F1", "4C88C488", "12345678"),4)) +
  #scale_linetype_manual(values=c("twodash", "twodash", "twodash", "solid")) +
  scale_x_continuous("Generations") +
  theme_classic() +
  #scale_color_brewer(palette="Spectral", direction = 1) +
  #scale_color_manual(values = c("#E5F5E0","#C7E9C0", "#A1D99B", "#74C476", "#41AB5D", "#238B45", "#006D2C", "#00441B","#2171B5", "#08519C", "#08306B")) + 
  #facet_grid(VESSEL ~ .) +
  #scale_color_manual(name="Type",values=c("black", "black", "#41ab5d")) +
  facet_wrap_paginate(~VESSEL, ncol = 2, nrow = 2, page = 3) +
  guides(colour = guide_legend(override.aes = list(size=2))) + 
  theme(legend.title = element_text(colour="black", size=15, face="bold"), axis.title.y = element_text(face="bold", size=14), axis.text.y = element_text(size=13), axis.title.x = element_text(face="bold", size=16), axis.text.x = element_text(size=15))

ggplot(alldata, aes(x=TIMEPOINT,y=(NORMALIZED_GFP_MEDIAN), colour=VESSEL)) +
  geom_point() +
  geom_line(aes(linetype=VESSEL), size=1.2) +
  #scale_y_continuous('Normalized fluorescence signal', limits=c(0,0.8)) +
  scale_linetype_manual(values=rep(c("solid", "dashed", "dotdash", "longdash", "twodash", "F1", "4C88C488", "12345678"),4)) +
  #scale_linetype_manual(values=c("twodash", "twodash", "twodash", "solid")) +
  scale_x_continuous("Generations") +
  theme_classic() +
  #scale_color_brewer(palette="Spectral", direction = 1) +
  #scale_color_manual(values = c("#E5F5E0","#C7E9C0", "#A1D99B", "#74C476", "#41AB5D", "#238B45", "#006D2C", "#00441B","#2171B5", "#08519C", "#08306B")) + 
  #facet_grid(VESSEL ~ .) +
  #scale_color_manual(name="Type",values=c("black", "black", "#41ab5d")) +
  facet_wrap_paginate(~VESSEL, ncol = 2, nrow = 2, page = 4) +
  guides(colour = guide_legend(override.aes = list(size=2))) + 
  theme(legend.title = element_text(colour="black", size=15, face="bold"), axis.title.y = element_text(face="bold", size=14), axis.text.y = element_text(size=13), axis.title.x = element_text(face="bold", size=16), axis.text.x = element_text(size=15))


ggplot(alldata, aes(x=TIMEPOINT,y=(NORMALIZED_GFP_MEDIAN), colour=VESSEL)) +
  geom_point() +
  geom_line(aes(linetype=VESSEL), size=1.2) +
  #scale_y_continuous('Normalized fluorescence signal', limits=c(0,0.8)) +
  scale_linetype_manual(values=rep(c("solid", "dashed", "dotdash", "longdash", "twodash", "F1", "4C88C488", "12345678"),4)) +
  #scale_linetype_manual(values=c("twodash", "twodash", "twodash", "solid")) +
  scale_x_continuous("Generations") +
  theme_classic() +
  #scale_color_brewer(palette="Spectral", direction = 1) +
  #scale_color_manual(values = c("#E5F5E0","#C7E9C0", "#A1D99B", "#74C476", "#41AB5D", "#238B45", "#006D2C", "#00441B","#2171B5", "#08519C", "#08306B")) + 
  #facet_grid(VESSEL ~ .) +
  #scale_color_manual(name="Type",values=c("black", "black", "#41ab5d")) +
  facet_wrap_paginate(~VESSEL, ncol = 2, nrow = 2, page = 5) +
  guides(colour = guide_legend(override.aes = list(size=2))) + 
  theme(legend.title = element_text(colour="black", size=15, face="bold"), axis.title.y = element_text(face="bold", size=14), axis.text.y = element_text(size=13), axis.title.x = element_text(face="bold", size=16), axis.text.x = element_text(size=15))

ggplot(alldata, aes(x=TIMEPOINT,y=(NORMALIZED_GFP_MEDIAN), colour=VESSEL)) +
  geom_point() +
  geom_line(aes(linetype=VESSEL), size=1.2) +
  #scale_y_continuous('Normalized fluorescence signal', limits=c(0,0.8)) +
  scale_linetype_manual(values=rep(c("solid", "dashed", "dotdash", "longdash", "twodash", "F1", "4C88C488", "12345678"),4)) +
  #scale_linetype_manual(values=c("twodash", "twodash", "twodash", "solid")) +
  scale_x_continuous("Generations") +
  theme_classic() +
  #scale_color_brewer(palette="Spectral", direction = 1) +
  #scale_color_manual(values = c("#E5F5E0","#C7E9C0", "#A1D99B", "#74C476", "#41AB5D", "#238B45", "#006D2C", "#00441B","#2171B5", "#08519C", "#08306B")) + 
  #facet_grid(VESSEL ~ .) +
  #scale_color_manual(name="Type",values=c("black", "black", "#41ab5d")) +
  facet_wrap_paginate(~VESSEL, ncol = 2, nrow = 2, page = 6) +
  guides(colour = guide_legend(override.aes = list(size=2))) + 
  theme(legend.title = element_text(colour="black", size=15, face="bold"), axis.title.y = element_text(face="bold", size=14), axis.text.y = element_text(size=13), axis.title.x = element_text(face="bold", size=16), axis.text.x = element_text(size=15))

ggplot(alldata, aes(x=TIMEPOINT,y=(NORMALIZED_GFP_MEDIAN), colour=VESSEL)) +
  geom_point() +
  geom_line(aes(linetype=VESSEL), size=1.2) +
  #scale_y_continuous('Normalized fluorescence signal', limits=c(0,0.8)) +
  scale_linetype_manual(values=rep(c("solid", "dashed", "dotdash", "longdash", "twodash", "F1", "4C88C488", "12345678"),4)) +
  #scale_linetype_manual(values=c("twodash", "twodash", "twodash", "solid")) +
  scale_x_continuous("Generations") +
  theme_classic() +
  #scale_color_brewer(palette="Spectral", direction = 1) +
  #scale_color_manual(values = c("#E5F5E0","#C7E9C0", "#A1D99B", "#74C476", "#41AB5D", "#238B45", "#006D2C", "#00441B","#2171B5", "#08519C", "#08306B")) + 
  #facet_grid(VESSEL ~ .) +
  #scale_color_manual(name="Type",values=c("black", "black", "#41ab5d")) +
  facet_wrap_paginate(~VESSEL, ncol = 2, nrow = 2, page = 7) +
  guides(colour = guide_legend(override.aes = list(size=2))) + 
  theme(legend.title = element_text(colour="black", size=15, face="bold"), axis.title.y = element_text(face="bold", size=14), axis.text.y = element_text(size=13), axis.title.x = element_text(face="bold", size=16), axis.text.x = element_text(size=15))

ggplot(alldata, aes(x=TIMEPOINT,y=(NORMALIZED_GFP_MEDIAN), colour=VESSEL)) +
  geom_point() +
  geom_line(aes(linetype=VESSEL), size=1.2) +
  #scale_y_continuous('Normalized fluorescence signal', limits=c(0,0.8)) +
  scale_linetype_manual(values=rep(c("solid", "dashed", "dotdash", "longdash", "twodash", "F1", "4C88C488", "12345678"),4)) +
  #scale_linetype_manual(values=c("twodash", "twodash", "twodash", "solid")) +
  scale_x_continuous("Generations") +
  theme_classic() +
  #scale_color_brewer(palette="Spectral", direction = 1) +
  #scale_color_manual(values = c("#E5F5E0","#C7E9C0", "#A1D99B", "#74C476", "#41AB5D", "#238B45", "#006D2C", "#00441B","#2171B5", "#08519C", "#08306B")) + 
  #facet_grid(VESSEL ~ .) +
  #scale_color_manual(name="Type",values=c("black", "black", "#41ab5d")) +
  facet_wrap_paginate(~VESSEL, ncol = 2, nrow = 2, page = 8) +
  guides(colour = guide_legend(override.aes = list(size=2))) + 
  theme(legend.title = element_text(colour="black", size=15, face="bold"), axis.title.y = element_text(face="bold", size=14), axis.text.y = element_text(size=13), axis.title.x = element_text(face="bold", size=16), axis.text.x = element_text(size=15))
```
```{r Median fl1 over time}

data$VESSEL <- as.factor(data$VESSEL)

alldata <- data

#median  fluor#
myplot <- ggplot(alldata, aes(x=TIMEPOINT,y=(FL1_MEDIAN), colour=VESSEL)) +
  geom_point() +
  geom_line(aes(linetype=VESSEL), size=1.2) +
  #scale_y_continuous('Normalized fluorescence signal', limits=c(0,0.8)) +
  #scale_linetype_manual(values=c("twodash", "twodash", "twodash", "solid")) +
  scale_x_continuous("Generations") +
  theme_classic() +
  #scale_color_brewer(palette="Spectral", direction = 1) +
  #scale_color_manual(values = c("#E5F5E0","#C7E9C0", "#A1D99B", "#74C476", "#41AB5D", "#238B45", "#006D2C", "#00441B","#2171B5", "#08519C", "#08306B")) + 
  #facet_grid(VESSEL ~ .) +
  #scale_color_manual(name="Type",values=c("black", "black", "#41ab5d")) +
  facet_wrap_paginate(~VESSEL, ncol = 2, nrow = 2, page = 1) +
  guides(colour = guide_legend(override.aes = list(size=2))) + 
  theme(legend.title = element_text(colour="black", size=15, face="bold"), axis.title.y = element_text(face="bold", size=14), axis.text.y = element_text(size=13), axis.title.x = element_text(face="bold", size=16), axis.text.x = element_text(size=15))


myplot #%+% subset(alldata, Nutrient %in% c("Glutamine"))
```


###PLOT NORMALIZED FLUOR###
#myplot <- ggplot(alldata, aes(x=TIMEPOINT,y=(NORMALIZED_GFP_MEDIAN), colour=POPULATION)) +
  geom_point() +
  geom_line(aes(linetype=POPULATION), size=1.2) +
  scale_y_continuous('Normalized fluorescence signal') +
  #scale_linetype_manual(values=c("twodash", "twodash", "twodash", "solid")) +
  scale_x_continuous("Generations") +
  theme_classic() +
  #scale_color_brewer(palette="Spectral", direction = 1) +
  #scale_color_manual(values = c("#E5F5E0","#C7E9C0", "#A1D99B", "#74C476", "#41AB5D", "#238B45", "#006D2C", "#00441B","#2171B5", "#08519C", "#08306B")) + 
  #facet_grid(Population ~ .) +
  #scale_color_manual(name="Type",values=c("black", "black", "#41ab5d")) +
  guides(colour = guide_legend(override.aes = list(size=2))) + 
  theme(legend.title = element_text(colour="black", size=15, face="bold"), axis.title.y = element_text(face="bold", size=14), axis.text.y = element_text(size=13), axis.title.x = element_text(face="bold", size=16), axis.text.x = element_text(size=15))

myplot #%+% subset(alldata, POPULATION %in% c("Genomic_Architecture_Mutant"))

```{r Proportion CNV cells}
###PLOT PROPORTION OF 2 copy CELLS###
myplot <- ggplot(alldata, aes(x=TIMEPOINT,y=(PopProp_2copy), colour=GENOTYPE)) +
  geom_point() +
  geom_line(size=1.2) +
  scale_y_continuous('Proportion of cells with 2 copies', limits=c(0,1)) +
  #scale_linetype_manual(values=c("twodash", "solid")) +
  scale_x_continuous("Generations") +
  theme_classic() +
  #scale_color_brewer(palette="Spectral", direction = 1) +
  #scale_color_manual(values = c("#E5F5E0","#C7E9C0", "#A1D99B", "#74C476", "#41AB5D", "#238B45", "#006D2C", "#00441B")) + 
  #facet_grid(Population ~ .) +
  #scale_color_manual(name="Type",values=c("black", "black", "#41ab5d")) +
  guides(colour = guide_legend(override.aes = list(size=2))) + 
  theme(legend.title = element_text(colour="black", size=15, face="bold"), axis.title.y = element_text(face="bold", size=14), axis.text.y = element_text(size=13), axis.title.x = element_text(face="bold", size=16), axis.text.x = element_text(size=15))

myplot #%+% subset(alldata, Nutrient %in% c("Glutamine") & Population %in% c("Experimental"))

###PLOT PROPORTION OF 3+ COPY CELLS###
myplot <- ggplot(alldata, aes(x=TIMEPOINT,y=(PopProp_3plus), colour=GENOTYPE)) +
  geom_point() +
  geom_line(size=1.2) +
  scale_y_continuous('Proportion of cells with 3+ copies', limits=c(0,1)) +
  #scale_linetype_manual(values=c("twodash", "solid")) +
  scale_x_continuous("Generations") +
  theme_classic() +
  #scale_color_brewer(palette="Spectral", direction = 1) +
  #scale_color_manual(values = c("#E5F5E0","#C7E9C0", "#A1D99B", "#74C476", "#41AB5D", "#238B45", "#006D2C", "#00441B")) + 
  #facet_grid(Population ~ .) +
  #scale_color_manual(name="Type",values=c("black", "black", "#41ab5d")) +
  guides(colour = guide_legend(override.aes = list(size=2))) + 
  theme(legend.title = element_text(colour="black", size=15, face="bold"), axis.title.y = element_text(face="bold", size=14), axis.text.y = element_text(size=13), axis.title.x = element_text(face="bold", size=16), axis.text.x = element_text(size=15))

myplot# %+% subset(alldata, Nutrient %in% c("Glutamine") & Population %in% c("Experimental"))

```

```{r FSC over time}
####FSC over time####
myplot <- ggplot(alldata, aes(x=TIMEPOINT,y=(FSC_MEDIAN), colour=GENOTYPE)) +
  geom_point() +
  geom_line(size=1.2) +
  scale_y_continuous('FSC') +
  #scale_linetype_manual(values=c("twodash", "solid")) +
  scale_x_continuous("Generations") +
  theme_classic() +
  #scale_color_brewer(palette="Spectral", direction = 1) +
  scale_color_manual(values = c('darkblue','#6baed6','blue', '#a1d99b','#74c476','#41ab5d','#238b45','#006d2c','#00441b', '#08306b', 'black')) + 
  #facet_grid(Population ~ .) +
  #scale_color_manual(name="Type",values=c("black", "black", "#41ab5d")) +
  guides(colour = guide_legend(override.aes = list(size=2))) + 
  theme(legend.title = element_text(colour="black", size=15, face="bold"), axis.title.y = element_text(face="bold", size=14), axis.text.y = element_text(size=13), axis.title.x = element_text(face="bold", size=16), axis.text.x = element_text(size=15))

myplot %+% subset(alldata, Nutrient %in% c("Glutamine"))

####FSC over time without subsetting####
myplot <- ggplot(alldata_unmasked, aes(x=TIMEPOINT,y=(FSC_MEDIAN), colour=GENOTYPE)) +
  geom_point() +
  geom_line(size=1.2) +
  scale_y_continuous('FSC all') +
  #scale_linetype_manual(values=c("twodash", "solid")) +
  scale_x_continuous("Generations") +
  theme_classic() +
  #scale_color_brewer(palette="Spectral", direction = 1) +
  scale_color_manual(values = c('darkblue','#6baed6','blue', '#a1d99b','#74c476','#41ab5d','#238b45','#006d2c','#00441b', '#08306b', 'black')) + 
  #facet_grid(Population ~ .) +
  #scale_color_manual(name="Type",values=c("black", "black", "#41ab5d")) +
  guides(colour = guide_legend(override.aes = list(size=2))) + 
  theme(legend.title = element_text(colour="black", size=15, face="bold"), axis.title.y = element_text(face="bold", size=14), axis.text.y = element_text(size=13), axis.title.x = element_text(face="bold", size=16), axis.text.x = element_text(size=15))

myplot %+% subset(alldata, Nutrient %in% c("Glutamine"))
```
```{r Barplots CNVs over time --- masked and unmasked}
new <- melt(alldata, id.vars = c("Generation", "Sample", "Nutrient"), measure.vars = c("PopProp_0copy","PopProp_1copy","PopProp_2copy","PopProp_3plus"))
new <- data.frame(Generation=as.character(new[,1]), Population=as.character(new[,2]), Nutrient=as.character(new[,3]), variable=as.character(new[,4]), value=as.numeric(new[,5])) 

new$Generation <- factor(new$Generation, levels = unique(alldata$Generation))
new$variable <- factor(new$variable, levels = c("PopProp_3plus", "PopProp_2copy", "PopProp_1copy", "PopProp_0copy"))

theplot <- ggplot(new, aes(Generation, value, fill=variable)) + 
  theme_classic() +
  theme(legend.position = "top", legend.title = element_blank(), axis.title.y = element_text(face="bold", size=14), axis.text.y = element_text(size=13), axis.title.x = element_text(face="bold", size=16)) + 
  geom_bar(colour = "black", position = "fill", stat = "identity", width = 0.9) +
  scale_y_continuous("Proportion", limits=c(0,1)) + 
  scale_fill_manual(values = c('#00441b', '#006d2c', '#a1d99b','#e5f5e0'))

theplot %+% subset(new, Population %in% c("gln_h_bc01"))


theplot %+% subset(new, Nutrient %in% c("Glutamine")) + 
  facet_grid(Population ~ .)


new <- melt(alldata_unmasked, id.vars = c("Generation", "Sample", "Nutrient"), measure.vars = c("PopProp_0copy","PopProp_1copy","PopProp_2copy","PopProp_3plus"))
new <- data.frame(Generation=as.character(new[,1]), Population=as.character(new[,2]), Nutrient=as.character(new[,3]), variable=as.character(new[,4]), value=as.numeric(new[,5])) 

new$Generation <- factor(new$Generation, levels = unique(alldata$Generation))
new$variable <- factor(new$variable, levels = c("PopProp_3plus", "PopProp_2copy", "PopProp_1copy", "PopProp_0copy"))

theplot <- ggplot(new, aes(Generation, value, fill=variable)) + 
  theme_classic() +
  theme(legend.position = "top", legend.title = element_blank(), axis.title.y = element_text(face="bold", size=14), axis.text.y = element_text(size=13), axis.title.x = element_text(face="bold", size=16)) + 
  geom_bar(colour = "black", position = "fill", stat = "identity", width = 0.9) +
  scale_y_continuous("Proportion", limits=c(0,1)) + 
  scale_fill_manual(values = c('#00441b', '#006d2c', '#a1d99b','#e5f5e0'))

theplot %+% subset(new, Population %in% c("gln_h_ctrl2"))


theplot %+% subset(new, Nutrient %in% c("Glutamine")) + 
  facet_grid(Population ~ .)
```

```{r}
#Load data.frame


ggplot(data2 %>% filter(VESSEL==5), aes(x = FL1.A, y = TIMEPOINT, fill = ..x.., height=..density..)) +
  geom_joy_gradient(scale = 1.25, rel_min_height = 0.02) +
  facet_grid(. ~ SAMPLE)
  scale_x_continuous("Normalized Fluorescence (a.u.)", limits=c(0.05,1.4), expand = c(0.01, 0), breaks = c(0.1, 0.7, 1.2)) +
  scale_y_discrete("Generations", expand = c(0.01, 0)) +
  facet_grid(. ~ SAMPLE) + 
  scale_fill_distiller(type = "seq", palette = 5, direction = 1, guide = "colourbar") +
  theme_bw() + 
  theme(text=element_text(size=11),legend.position = 'none', legend.title = element_blank(), axis.title.y = element_text(face="bold", size=12), axis.text.y = element_text(size=10), axis.title.x = element_text(face="bold", size=12), axis.text.x = element_text(size=10))
```