---
title: "LTEE_GAP1_CNV_Variant_Analysis"
author: '`r Sys.info()[7]`'
date: '`r Sys.Date()`'
output:
  html_document:
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
---

**Experiment overview**

Write a detailed description of your experiment here including the goal of the analysis and your interpretation of the results.   
If you still see this text it means that you have not described the experiment and whatever follows is meaningless.
###############################

> This code is designed for use with the Accuri flow cytometer, which is equiped with the following lasers and filters

* Blue laser (488 nm)
  + FL1 filter = 514/20nm   GFP
  + FL3 filter = 575/25nm   YFP

* Yellow/green laser (552 nm)
  + FL2 filter = 610/20nm   mCherry, dtomato
  + FL4 filter = 586/15nm   DsRed
  
**Requirements**  

**Output**  



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Step 1: Load relevant libraries 

```{r}
# This is a function that just makes sure you have a package, or installs it for you without prompting

requireInstall <- function(packageName,isBioconductor=F) {
  if ( !try(require(packageName,character.only=T)) ) {
    print(paste0("You don't have ",packageName," accessible, ",
      "I'm gonna install it"))
    if (isBioconductor) {
      source("http://bioconductor.org/biocLite.R")                        
      biocLite(packageName)                                                 
    } else {
      install.packages("packageName", repos = "http://cran.us.r-project.org")
    }
  }
  return(1)
}

#Load libraries
requireInstall("tidyverse")
requireInstall("flowCore",isBioconductor=T)
requireInstall("flowViz",isBioconductor=T)
requireInstall("ggcyto",isBioconductor=T)
requireInstall("devtools",isBioconductor=T)
requireInstall("ggpubr",isBioconductor=T)
devtools::install_github("guiastrennec/ggplus")
devtools::install_github("kassambara/ggpubr")
library("ggplus")
library("ggpubr")
```

```{r}
#Set working directory to the folder in which you have stored your .fcs files
#Read in all the fcs files in the directory.

path.data <-"~/Google Drive/MiniStatRun_10_2018/"

list.folders <- c("LTEE_mCitrine_GAP1_Variants_T00", "LTEE_mCitrine_GAP1_Variants_T06", "LTEE_mCitrine_GAP1_Variants_T07", "LTEE_mCitrine_GAP1_Variants_T08.1", "LTEE_mCitrine_GAP1_Variants_T08.2", "LTEE_mCitrine_GAP1_Variants_T08.3", "LTEE_mCitrine_GAP1_Variants_T11.1", "LTEE_mCitrine_GAP1_Variants_T11.2", "LTEE_mCitrine_GAP1_Variants_T13.1", "LTEE_mCitrine_GAP1_Variants_T13.2", "LTEE_mCitrine_GAP1_Variants_T14", "LTEE_mCitrine_GAP1_Variants_T15", "LTEE_mCitrine_GAP1_Variants_T18")
  
flowData <- read.flowSet(path = paste(path.data, list.folders[1],"/", sep=""), pattern=".fcs", alter.names = TRUE)

sample.sheet <- read.csv(paste(path.data,"Samplesheet_",list.folders[1],".csv", sep=""))

#Adds a sample sheet data to the pData of the flowset

sampleNames(flowData) <- paste(gsub(" ","_",sample.sheet$Strain),"_T",gsub(" ","_",sample.sheet$Timepoint),"_",sub(" ","_",sample.sheet$Well), sep="")
pData(flowData)$Strain <- gsub(" ","_",sample.sheet$Strain)
pData(flowData)$Genotype <- gsub(" ","_",sample.sheet$Genotype)
pData(flowData)$Classification <- gsub(" ","_",sample.sheet$Classification)
pData(flowData)$Vessel <- gsub(" ","_",sample.sheet$Vessel)
pData(flowData)$Media <- gsub(" ","_",sample.sheet$Media)
pData(flowData)$Ploidy <- gsub(" ","_",sample.sheet$Ploidy)
pData(flowData)$Well <- gsub(" ","_",sample.sheet$Well)
pData(flowData)$Experiment <- gsub(" ","_",sample.sheet$Experiment)
pData(flowData)$Timepoint <- gsub(" ","_",sample.sheet$Timepoint)
pData(flowData)$Run <- list.folders[1]

#Need to go throw list 1 by 1 troubnleshooting errors
for(i in 2:length(list.folders)) {
  flowData.temp <- read.flowSet(path = paste(path.data,list.folders[i],"/", sep=""), pattern=".fcs", alter.names = TRUE)

  sample.sheet <- read.csv(paste(path.data,"Samplesheet_",list.folders[i],".csv", sep=""))

  #Adds a sample sheet data to the pData of the flowset
  sampleNames(flowData.temp) <- paste(gsub(" ","_",sample.sheet$Strain),"_T",gsub(" ","_",sample.sheet$Timepoint),"_",sub(" ","_",sample.sheet$Well), sep="")
  pData(flowData.temp)$Strain <- gsub(" ","_",sample.sheet$Strain)
  pData(flowData.temp)$Genotype <- gsub(" ","_",sample.sheet$Genotype)
  pData(flowData.temp)$Classification <- gsub(" ","_",sample.sheet$Classification)
  pData(flowData.temp)$Vessel <- gsub(" ","_",sample.sheet$Vessel)
  pData(flowData.temp)$Media <- gsub(" ","_",sample.sheet$Media)
  pData(flowData.temp)$Ploidy <- gsub(" ","_",sample.sheet$Ploidy)
  pData(flowData.temp)$Well <- gsub(" ","_",sample.sheet$Well)
  pData(flowData.temp)$Experiment <- gsub(" ","_",sample.sheet$Experiment)
  pData(flowData.temp)$Timepoint <- gsub(" ","_",sample.sheet$Timepoint)
  pData(flowData.temp)$Run <- list.folders[i]

  flowData <- rbind2(flowData, flowData.temp)
}
```

```{r flowSet summaries}
#Check how many cells were counted in each fcs file
total <- fsApply(flowData, each_col, length) #total counts per sample
print(total)
#Print the medians of data values for each measurement
fsApply(flowData, each_col, median)

#combine all flowSets into a single flowset
samples.num <- length(flowData)
print(samples.num)
```
#Step 3: apply filters to data and generate plots showing the effect on filtering

#Step 4: Data analysis


#diagnostic values can be defined for plotting purposes
```{r Definition of diagnostic values}
#define critical values that can superimposed on plots for easy visual comparison

gfp.bg <- 3.9  #a background value for GFP
gfp.wt <- 5.9 #a value for wildtype GFP expression
haploid.fsc <- 6e5 #an empirical value for forward scatter for haploids
diploid.fsc <- 7e5 #an empirical value for forward scatter for diploids
gfp.norm <- 0.935 #an empricial value for gfp expression normalized by forward scatter

```


#Transfer data into a dataframe and produce summary stats
```{r}
#Temp
filterData <-flowData


#Summary Statistics of Normalized Filtered Data
#Move filtered data into a dataframe
filter.data <- data.frame(FSC.A=NA,SSC.A=NA,FL1.A=NA, SAMPLE=NA, VESSEL=NA,STRAIN=NA,TIMEPOINT=NA)
stats.data <- data.frame(VESSEL=NA,STRAIN=NA,COUNT=NA,MEDIAN=NA,MEAN=NA,SD=NA,VAR=NA)

for(i in 1:length(filterData)){
  fsc.a <- exprs(filterData[[i,1]])
  ssc.a <- exprs(filterData[[i,2]])
  fl1.a <- exprs(filterData[[i,3]])
  #fl2.a <- exprs(filterData[[i,4]])
  #fl3.a <- exprs(filterData[[i,5]])
  #fl4.a <- exprs(filterData[[i,6]])
  sample <- sampleNames(filterData[i])
  vessel <- pData(filterData)$Vessel[i]
  strain <- pData(filterData)$Strain[i]
  timepoint <- pData(filterData)$Timepoint[i]
  

  filter.data<-rbind(filter.data,cbind(FSC.A=fsc.a,SSC.A=ssc.a,FL1.A=fl1.a,SAMPLE=sample,VESSEL=vessel,STRAIN=strain,TIMEPOINT=timepoint))
  stats.data<-(rbind(stats.data,cbind(VESSEL=vessel,STRAIN=strain,COUNT=length(fsc.a),MEDIAN=median(fl1.a/fsc.a),MEAN=mean(fl1.a/fsc.a),SD=sd(fl1.a/fsc.a),VAR=as.numeric(var(fl1.a/fsc.a)))))
} 

#Cleans up DataFrames
filter.data<-filter.data[2:nrow(filter.data),]
  
stats.data<-stats.data[2:nrow(stats.data),]
```

```{r}
#Saves Data
save(path.data, list.folders, total, samples.num, flowData, filterData, filter.data, stats.data, file = paste(path.data, "data_DF.RData", sep=""))
```

```{r}
#Loads Data
load(file = paste(path.data, "data_DF.RData", sep=""))
```

##Extract data from fcs files to generate statistics and boxplots
```{r Data extraction and plotting}
#Temp
filterData <-flowData

###GLUTAMINE###
#record summary statistics for each sample in a matrix named summary.stats 
summary.stats <- matrix(data = NA, nrow = length(filterData), ncol = 9, dimnames = list(sampleNames(filterData),c("FSC_median","FSC_mean", "FSC_sd","FL1_median", "FL1_mean","FL1_sd","normalizedGFP_median", "normalizedGFP_mean", "normalizedGFP_sd")))

#use the sample containing the minimum number of points after filtering for doublets and debris to define the number of data points retained for all samples
sample.size <- min(fsApply(filterData, each_col, length))  

print(sample.size)

comparison.FSC <- matrix(data = NA, nrow = sample.size, ncol = length(filterData), byrow = FALSE,dimnames = NULL)
comparison.FL1 <- matrix(data = NA, nrow = sample.size, ncol = length(filterData), byrow = FALSE,dimnames = NULL)
comparison.FL1NormFsc <- matrix(data = NA, nrow = sample.size, ncol = length(filterData), byrow = FALSE,dimnames = NULL)

#for each sample plot a histogram of the normalized data, raw FSC and raw GFP per row
par(mfrow=c(1,2), mar=c(5.1,2.1,2.1,2.1), oma=c(1.5,2,1,1))

#extract data from flowFrames to plot histograms of values and record summary statistics
for (i in 1:length(filterData)){
 
  temp <- exprs(filterData[[i]]) #exprs() extracts a matrix of the values from the flowframe
 

  ##########################################
  #record summary statistics for the sample#
  ##########################################
  
  #FSC
  summary.stats[i,1] <- median(temp[,1]) 
  summary.stats[i,2] <-mean(temp[,1])  
  summary.stats[i,3] <- sd(temp[,1])
  #FL1
  summary.stats[i,4] <- median(temp[,3])
  summary.stats[i,5] <-mean(temp[,3])  
  summary.stats[i,6] <- sd(temp[,3])
  #FL1 (GFP) divided by FSC
  summary.stats[i,7] <- median(temp[,3]/temp[,1])
  summary.stats[i,8] <-mean(temp[,3]/temp[,1])  
  summary.stats[i,9] <- sd(temp[,3]/temp[,1])
  
  ##############################################
  #plot histograms of the channels of interest##
  ##############################################

  ###############
  #Green channel#
  ###############
  
  #FL1 (GFP)
  hist(log10(temp[,3]), br=1000, xlab = "log10(FL1)", main = "FL1") 
  abline(v=gfp.bg, col="yellow", lty=2, lwd=2)
  abline(v=gfp.wt, col="green", lty=2, lwd=2) 
  legend("topleft",  legend=paste("median FL1 = ",round(median(temp[,3]), digits=4),sep=""))

  #GFP divided by FSC
  hist(temp[,3]/temp[,1], br=500, xlab = "FL1/FSC", main = "FL1/FSC") 
  abline(v=gfp.norm, col="green", lty=2, lwd=2 )
  legend("topleft",  legend=paste("median GFP / FSC=",round(median(temp[,3]/temp[,1]), digits=4),sep=""))
  
  mtext(sampleNames(filterData[i]), outer = TRUE, cex = 1.0)
  
  ###############
  #Other#########
  ###############

    #FSC
  hist(log10(temp[,1]), br=500, xlab = "log10(FSC)", main = "FSC", xlim=c(4,8)) 
  abline(v=haploid.fsc, col="blue", lty=2, lwd=2)
  abline(v=diploid.fsc, col="grey", lty=2, lwd=2)
  legend("topleft",  legend=paste("median FSC=",round(median(temp[,1]), digits=4),sep=""))
  
  mtext(sampleNames(filterData[i]), outer = TRUE, cex = 1.0)

print("-------------------------------------------------------")
print("-----------------------------------")
print("----------------------")

  ############################################################
  #keep the data set for generating boxplots comparing values#
  ############################################################
  
  #Note that the amount of data kept for each sample is defined by the lowest count among all the samples.
  comparison.FSC[1:sample.size,i] <- temp[1:sample.size,1] #FSC
  comparison.FL1[1:sample.size,i] <- temp[1:sample.size,3] #FL1 (GFP)
  comparison.FL1NormFsc[1:sample.size,i] <- temp[1:sample.size,3]/temp[1:sample.size,1] #GFP/FSC
}

```


##Overview of data distributions
```{r Overall data distributions}
par(mar=c(8.1,4.1,4.1,2.1)) #create more space at lower margin

###GLUTAMINE###
boxplot(comparison.FSC, names=sampleNames(filterData), notch = TRUE, col = "gray", ylab="FSC", cex.axis=0.5,las=2, outline=F)
abline(h=haploid.fsc, lty=2, col=2)
abline(h=diploid.fsc, lty=2, col=3)

boxplot(comparison.FL1, names=sampleNames(filterData), notch = TRUE, col = "lightgreen", ylab="FL1", cex.axis=0.5,las=2, outline=F)
abline(h=gfp.bg ,lty=2, lwd=3, col="yellow")
abline(h=gfp.wt, lty = 2, lwd=3, col="green")

boxplot(comparison.FL1NormFsc, names=sampleNames(filterData), notch = TRUE, col = "green", ylab="FL1/FSC", cex.axis=0.5,las=2, outline=F)
abline(h=gfp.norm, lty=2, lwd=3, col="blue")

par(mar=c(5.1,4.1,4.1,2.1)) #reset margins to default

#generate summary tables containing all the recorded statistics
print(summary.stats)
summary.stats.gln <- as.data.frame(summary.stats)

```

```{r}
#Saves Data
save(path.data, list.folders, total, samples.num, flowData, filterData, summary.stats, sample.size, file = paste(path.data, "data_M.RData", sep=""))
```

```{r}
#Loads Data
load(file = paste(path.data, "data_M.RData", sep=""))
```


