---
title: "Gresham Lab Flow Core Guide"
author: '`r Sys.info()[7]`'
date: '`r Sys.Date()`'
output:
  html_document:
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
---


**Experiment overview**

THIS VERSION HAS BEEN EDITED TO REMOVE ALL ANALYSIS OF UREA LIMITED POPULATIONS (since the barcoded library can't grow in urea)

This markdown is modified for the experiment EE_GA_glnurmix_barcode17, which is involves the experimental evolution of haploid S. cerevisiae using miniature chemostats. 11 populations have glutamine media (0 copy control DGY1, 1 copy control DGY500, 2 copy control DGY1315, 8 experimental L139 in DGY1706), 11 populations have urea media (0 copy control DGY1, 1 copy control DGY500, 2 copy control DGY1315, 8 experimental L139 in DGY1706), and 10 populations have media with a mix of glutamine and urea (0 copy control DGY1, 1 copy control DGY500, 2 copy control DGY1315, 7 experimental L139 in DGY1706).

This particular markdown is for sample 33, collected 17 Oct 2017, about 265 generations.
###############################

> This code is designed for use with the Accuri flow cytometer, which is equiped with the following lasers and filters

* Blue laser (488 nm)
  + FL1 filter = 514/20nm   GFP
  + FL3 filter = 575/25nm   YFP

* Yellow/green laser (552 nm)
  + FL2 filter = 610/20nm   mCherry, dtomato
  + FL4 filter = 586/15nm   DsRed
  
  

**Requirements**  
In order to run this code you need:  
  + to predefine your gates using the **gating.R** script  
  + the **gates.Rdata** workspace, which contains the gates used in this script  
  + the path of the directory(ies), given the variable names **dir1**, **dir2**... that contain .fcs files named A01.fcs, A02.fcs, A03.fcs...  
  + a tab delimited sample sheet in each directory with the following rows: <Well>	<Strain>	<Genotype>	<Ploidy>	<Media>	<Experiment>  
  + the variable names are changed in chunk 2 named "Variable Names"    




**Output**  
This script generates a summary of results followed by quality control plots.  



#Step 1: Load relevant libraries 
```{r Libraries, eval=TRUE}
# This is a function that just makes sure you have a package, or installs it for you without prompting

requireInstall <- function(packageName,isBioconductor=F) {
  if ( !try(require(packageName,character.only=T)) ) {
    print(paste0("You don't have ",packageName," accessible, ",
      "I'm gonna install it"))
    if (isBioconductor) {
      source("http://bioconductor.org/biocLite.R")                        
      biocLite(packageName)                                                 
    } else {
      install.packages("packageName", repos = "http://cran.us.r-project.org")
    }
  }
  return(1)
}

#Load libraries
requireInstall("flowCore",isBioconductor=T)
requireInstall("flowViz",isBioconductor=T)
requireInstall("flowStats")
requireInstall("Hmisc")
requireInstall("reshape2")
requireInstall("ggplot2")
requireInstall("flowWorkspace")
requireInstall("ggcyto", isBioconductor=T)
requireInstall("gridExtra")
requireInstall("stringr")


```

#Step 2: Read in .fcs files, an Rdata file containing the gates sample sheet(s) that contains four columns with 
* column1 = Well
* column2 = Strain
* column3 = Staining
* column4 = Media
* column5 = Userdefined

```{r}
#Set working directory to the folder in which you have stored your .fcs files
#Read in all the fcs files in the directory.

path.data <-"~/Google Drive/MiniStatRun_10_2018/"

list.folders <- c("LTEE_mCitrine_GAP1_Variants_T00", "LTEE_mCitrine_GAP1_Variants_T06", "LTEE_mCitrine_GAP1_Variants_T07", "LTEE_mCitrine_GAP1_Variants_T08.1", "LTEE_mCitrine_GAP1_Variants_T08.2", "LTEE_mCitrine_GAP1_Variants_T08.3", "LTEE_mCitrine_GAP1_Variants_T11.1", "LTEE_mCitrine_GAP1_Variants_T11.2", "LTEE_mCitrine_GAP1_Variants_T13.1", "LTEE_mCitrine_GAP1_Variants_T13.2", "LTEE_mCitrine_GAP1_Variants_T14", "LTEE_mCitrine_GAP1_Variants_T15", "LTEE_mCitrine_GAP1_Variants_T18")
  
flowData <- read.flowSet(path = paste(path.data, list.folders[1],"/", sep=""), pattern=".fcs", alter.names = TRUE)

sample.sheet <- read.csv(paste(path.data,"Samplesheet_",list.folders[1],".csv", sep=""))

#Adds a sample sheet data to the pData of the flowset

sampleNames(flowData) <- paste(gsub(" ","_",sample.sheet$Strain),"_T",gsub(" ","_",sample.sheet$Timepoint),"_",sub(" ","_",sample.sheet$Well), sep="")
pData(flowData)$Strain <- gsub(" ","_",sample.sheet$Strain)
pData(flowData)$Genotype <- gsub(" ","_",sample.sheet$Genotype)
pData(flowData)$Classification <- gsub(" ","_",sample.sheet$Classification)
pData(flowData)$Vessel <- gsub(" ","_",sample.sheet$Vessel)
pData(flowData)$Media <- gsub(" ","_",sample.sheet$Media)
pData(flowData)$Ploidy <- gsub(" ","_",sample.sheet$Ploidy)
pData(flowData)$Well <- gsub(" ","_",sample.sheet$Well)
pData(flowData)$Experiment <- gsub(" ","_",sample.sheet$Experiment)
pData(flowData)$Timepoint <- gsub(" ","_",sample.sheet$Timepoint)
pData(flowData)$Run <- list.folders[1]

#Need to go throw list 1 by 1 troubnleshooting errors
for(i in 2:length(list.folders)) {
  flowData.temp <- read.flowSet(path = paste(path.data,list.folders[i],"/", sep=""), pattern=".fcs", alter.names = TRUE)

  sample.sheet <- read.csv(paste(path.data,"Samplesheet_",list.folders[i],".csv", sep=""))

  #Adds a sample sheet data to the pData of the flowset
  sampleNames(flowData.temp) <- paste(gsub(" ","_",sample.sheet$Strain),"_T",gsub(" ","_",sample.sheet$Timepoint),"_",sub(" ","_",sample.sheet$Well), sep="")
  pData(flowData.temp)$Strain <- gsub(" ","_",sample.sheet$Strain)
  pData(flowData.temp)$Genotype <- gsub(" ","_",sample.sheet$Genotype)
  pData(flowData.temp)$Classification <- gsub(" ","_",sample.sheet$Classification)
  pData(flowData.temp)$Vessel <- gsub(" ","_",sample.sheet$Vessel)
  pData(flowData.temp)$Media <- gsub(" ","_",sample.sheet$Media)
  pData(flowData.temp)$Ploidy <- gsub(" ","_",sample.sheet$Ploidy)
  pData(flowData.temp)$Well <- gsub(" ","_",sample.sheet$Well)
  pData(flowData.temp)$Experiment <- gsub(" ","_",sample.sheet$Experiment)
  pData(flowData.temp)$Timepoint <- gsub(" ","_",sample.sheet$Timepoint)
  pData(flowData.temp)$Run <- list.folders[i]

  flowData <- rbind2(flowData, flowData.temp)
}
```

```{r flowSet summaries}
#Check how many cells were counted in each fcs file
total <- fsApply(flowData, each_col, length)[1:length(flowData)] #total counts per sample
print(total)
#Print the medians of data values for each measurement
fsApply(flowData, each_col, median)[1:length(flowData)]

#combine all flowSets into a single flowset
samples.num <- length(flowData)
print(samples.num)
```

#Subsetting the experiments data
```{r Subset the Data}
#filt <- "3"
#filterData <- subset(flowData, pData(flowData)$Vessel == filt) #Sperates based on Vessel
#filterData <- subset(flowData, pData(flowData)$Timepoint == filt)

filterData <- flowData
```

```{r subsetted flowSet summaries}
#Check how many cells were counted in each fcs file
total <- fsApply(filterData, each_col, length)[1:length(filterData)] #total counts per sample
print(total)
#Print the medians of data values for each measurement
fsApply(filterData, each_col, median)[1:length(filterData)]

#combine all flowSets into a single flowset
samples.num <- length(filterData)
print(samples.num)
```


#Step 3: apply filters to data and generate plots showing the effect on filtering

#Gating
```{r Application of Gates}
##Subset the data by applying sequential gates##

#Load Gates
load(file = "gates_LTEE_GAP1_Var.RData")

#apply doublet gate to ALL SAMPLES
filterData.singlets <- Subset(filterData, pg.singlets) 
fsApply(filterData.singlets, each_col, length)[1:samples.num]
singlets <- fsApply(filterData.singlets, each_col, length)[1:samples.num]
barplot(singlets/total, ylim=c(0,1), ylab = "Proportion singlet cells", las=2, cex.names = 0.5, names.arg=sampleNames(filterData))


####GLUTAMINE GATES####
#apply debris gates
filtered.debris <- Subset(filterData, pg.nondebris) 
fsApply(filtered.debris, each_col, length)[1:samples.num]
non.debris <- fsApply(filtered.debris, each_col, length)[1:samples.num]
barplot(non.debris/total, ylim=c(0,1), ylab = "Proportion nondebris cells", las=2, cex.names = 0.5, names.arg=str_sub(sampleNames(filterData)))

#this gate defines nongfp cells
gfp.neg <- Subset(filterData, gln.zero) 
fsApply(gfp.neg, each_col, length)[1:samples.num]
non.gfp <- fsApply(gfp.neg, each_col, length)[1:samples.num]
barplot(non.gfp/non.debris, ylim=c(0,1), ylab = "Proportion cells with no GFP", las=2, cex.names = 0.5, names.arg=str_sub(sampleNames(filterData)))

#this gate defines one copy gfp cells
gfp.one <- Subset(filterData, gln.one) 
fsApply(gfp.one, each_col, length)[1:samples.num]
gfp.one.cells <- fsApply(gfp.one, each_col, length)[1:samples.num]
barplot(gfp.one.cells/non.debris, ylim=c(0,1), ylab = "Proportion cells with 1 copy GFP", las=2, cex.names = 0.5, names.arg=str_sub(sampleNames(filterData)))

#this gate defines two copy gfp cells
gfp.two <- Subset(filterData, gln.two) 
fsApply(gfp.two, each_col, length)[1:samples.num]
gfp.two.cells <- fsApply(gfp.two, each_col, length)[1:samples.num]
  barplot(gfp.two.cells/non.debris, ylim=c(0,1), ylab = "Proportion cells with 2 copy GFP", las=2, cex.names = 0.5, names.arg=str_sub(sampleNames(filterData)))

#this gate defines thre copy plus gfp cells
gfp.thr <- Subset(filterData, gln.three) 
fsApply(gfp.thr, each_col, length)[1:samples.num]
gfp.thr.cells <- fsApply(gfp.thr, each_col, length)[1:samples.num]
barplot(gfp.thr.cells/non.debris, ylim=c(0,1), ylab = "Proportion cells with 3 copy+ GFP", las=2, cex.names = 0.5, names.arg=str_sub(sampleNames(filterData)))
```
#Step 4: Data analysis

##diagnostic values can be defined for plotting purposes
```{r Definition of diagnostic values}
#define critical values that can superimposed on plots for easy visual comparison

gfp.bg <- 3.9  #a background value for GFP
gfp.wt <- 5.9 #a value for wildtype GFP expression
haploid.fsc <- 6e5 #an empirical value for forward scatter for haploids
diploid.fsc <- 7e5 #an empirical value for forward scatter for diploids
gfp.norm <- 0.935 #an empricial value for gfp expression normalized by forward scatter

```

##Extract data from fcs files to generate statistics and boxplots
```{r Data extraction and plotting}
####GLUTAMINE####
#record summary statistics for each sample in a matrix named summary.stats 
summary.stats <- matrix(data = NA, nrow = length(filterData), ncol = 9, dimnames = list(sampleNames(filterData),c("FSC_median","FSC_mean", "FSC_sd","FL1_median", "FL1_mean","FL1_sd","normalizedGFP_median", "normalizedGFP_mean", "normalizedGFP_sd")))

#use the sample containing the minimum number of points after filtering for doublets and debris to define the number of data points retained for all samples
sample.size <- min(fsApply(filterData, each_col, length))  

print(sample.size)

comparison.FSC <- matrix(data = NA, nrow = sample.size, ncol = length(filterData), byrow = FALSE,dimnames = NULL)
comparison.FL1 <- matrix(data = NA, nrow = sample.size, ncol = length(filterData), byrow = FALSE,dimnames = NULL)
comparison.FL1NormFsc <- matrix(data = NA, nrow = sample.size, ncol = length(filterData), byrow = FALSE,dimnames = NULL)

#for each sample plot a histogram of the normalized data, raw FSC and raw GFP per row
par(mfrow=c(1,2), mar=c(5.1,2.1,2.1,2.1), oma=c(1.5,2,1,1))

#extract data from flowFrames to plot histograms of values and record summary statistics
for (i in 1:length(filterData)){
 
  temp <- exprs(filterData[[i]]) #exprs() extracts a matrix of the values from the flowframe
 

  ##########################################
  #record summary statistics for the sample#
  ##########################################
  
  #FSC
  summary.stats[i,1] <- median(temp[,1]) 
  summary.stats[i,2] <-mean(temp[,1])  
  summary.stats[i,3] <- sd(temp[,1])
  #FL1
  summary.stats[i,4] <- median(temp[,3])
  summary.stats[i,5] <-mean(temp[,3])  
  summary.stats[i,6] <- sd(temp[,3])
  #FL1 (GFP) divided by FSC
  summary.stats[i,7] <- median(temp[,3]/temp[,1])
  summary.stats[i,8] <-mean(temp[,3]/temp[,1])  
  summary.stats[i,9] <- sd(temp[,3]/temp[,1])
  
  ##############################################
  #plot histograms of the channels of interest##
  ##############################################

  ###############
  #Green channel#
  ###############
  
  #FL1 (GFP)
  hist(log10(temp[,3]), br=1000, xlab = "log10(FL1)", main = "FL1") 
  abline(v=gfp.bg, col="yellow", lty=2, lwd=2)
  abline(v=gfp.wt, col="green", lty=2, lwd=2) 
  legend("topleft",  legend=paste("median FL1 = ",round(median(temp[,3]), digits=4),sep=""))

  #GFP divided by FSC
  hist(temp[,3]/temp[,1], br=500, xlab = "FL1/FSC", main = "FL1/FSC") 
  abline(v=gfp.norm, col="green", lty=2, lwd=2 )
  legend("topleft",  legend=paste("median GFP / FSC=",round(median(temp[,3]/temp[,1]), digits=4),sep=""))
  
  mtext(str_sub(sampleNames(flowData[i]), 1, 11), outer = TRUE, cex = 1.0)
  
  ###############
  #Other#########
  ###############

    #FSC
  hist(log10(temp[,1]), br=500, xlab = "log10(FSC)", main = "FSC", xlim=c(4,8)) 
  abline(v=haploid.fsc, col="blue", lty=2, lwd=2)
  abline(v=diploid.fsc, col="grey", lty=2, lwd=2)
  legend("topleft",  legend=paste("median FSC=",round(median(temp[,1]), digits=4),sep=""))
  
  mtext(str_sub(sampleNames(flowData[i]), 1,11), outer = TRUE, cex = 1.0)

print("-------------------------------------------------------")
print("-----------------------------------")
print("----------------------")

  ############################################################
  #keep the data set for generating boxplots comparing values#
  ############################################################
  
  #Note that the amount of data kept for each sample is defined by the lowest count among all the samples.
  comparison.FSC[1:sample.size,i] <- temp[1:sample.size,1] #FSC
  comparison.FL1[1:sample.size,i] <- temp[1:sample.size,3] #FL1 (GFP)
  comparison.FL1NormFsc[1:sample.size,i] <- temp[1:sample.size,3]/temp[1:sample.size,1] #GFP/FSC
}


par(mfrow=c(1,1)) #change number of plots per row back to standard

```


##Overview of data distributions
```{r Overall data distributions}
par(mar=c(8.1,4.1,4.1,2.1)) #create more space at lower margin

####GLUTAMINE####
boxplot(comparison.FSC, names=str_sub(sampleNames(flowData),1,11), notch = TRUE, col = "gray", ylab="FSC", cex.axis=0.5,las=2, outline=F)
abline(h=haploid.fsc, lty=2, col=2)
abline(h=diploid.fsc, lty=2, col=3)

boxplot(comparison.FL1, names=str_sub(sampleNames(flowData),1,11), notch = TRUE, col = "lightgreen", ylab="FL1", cex.axis=0.5,las=2, outline=F)
abline(h=gfp.bg ,lty=2, lwd=3, col="yellow")
abline(h=gfp.wt, lty = 2, lwd=3, col="green")

boxplot(comparison.FL1NormFsc, names=str_sub(sampleNames(flowData),1,11), notch = TRUE, col = "green", ylab="FL1/FSC", cex.axis=0.5,las=2, outline=F)
abline(h=gfp.norm, lty=2, lwd=3, col="blue")


par(mar=c(5.1,4.1,4.1,2.1)) #reset margins to default

#generate summary tables containing all the recorded statistics
print(summary.stats)
summary.stats <- as.data.frame(summary.stats)


```


##Quantitation of relative FL1 signal
```{r Quantitation of relative FL1 signal}

###GLUTAMINE###
baseline.FL1 <- summary.stats$FL1_median[4] #MUST REPLACE WITH YOUR 1 COPY CONTROL!!
barplot(summary.stats$FL1_median/baseline.FL1, ylim = c(0,2), ylab="Relative FL1 median expression", las=2, cex.names = 0.5, names.arg=str_sub(sampleNames(flowData),1, 11))

```

##Quantitation of forward scatter
```{r Quantitation of forward scatter}
#use 0 copy control
###GLUTAMINE###
baseline.FSC <- summary.stats$FSC_median[1]
barplot(summary.stats$FSC_median/baseline.FSC, ylab="Relative median FSC", las=2, cex.names = 0.5, names.arg=str_sub(sampleNames(flowData),1,11))


```

##Population composition
```{r}
###GLUTAMINE###
pop.composition <- rbind(non.gfp/non.debris,gfp.one.cells/non.debris,gfp.two.cells/non.debris, gfp.thr.cells/non.debris)
barplot(pop.composition, ylab="Proportion of population", legend=c("No GFP", "One copy GFP", "Two copy GFP", "Three+ copy GFP"),las=2, ylim=c(0,1), col=c("black", "grey", "light green", "dark green"), cex.names = 0.5,names.arg=str_sub(sampleNames(flowData),1,11))

```

#Step 5: Quality control

##Gates # Currently only works for a singal sample at a time
```{r Gates}
#Singlets gate
xyplot(FSC.A~FSC.H, data=flowData[1], xlim=c(0,3e6), ylim=c(0,3e6), filter=pg.singlets,  smooth=F, xbin=1024, stat=T, pos=0.5, abs=T, main = "First flowset - singlets gate")

###Glutamine###
#Debris gate
xyplot(SSC.A ~ FSC.A, data=flowData[1], displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=pg.nondebris, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T,  main = "Glutamine - nondebris gate")

#Non-fluorescent population gate
xyplot(FL1.A~FSC.A,data=filterData[1], displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=gln.zero, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T,  main = "Glutamine - non GFP gate")

#ONe copy Fluorescent population gate
xyplot(FL1.A~FSC.A,data=filterData[2], displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=gln.one, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T, main = "Glutamine - One copy GFP gate")

#Two copy fluorescent population gate
xyplot(FL1.A~FSC.A,data=filterData[3], displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=gln.two, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T, main = "Glutamine - Two copy GFP gate")

#Three copy plus fluorescent population gate
xyplot(FL1.A~FSC.A,data=filterData[3], displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=gln.three, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T, main = "Glutamine - Three+ copy GFP gate")

```


##Data transformation for visualization
```{r}
#In order to look at QC plots the data is transformed using the logicle transform, which is a log transform for high values that transitions to a linear transformation near zero values 

#This is simply for visualization purposes

lgcl <- logicleTransform(w = 0.5, t= 10000, m=4.5) #the parameters w,t, and m define the transformation parameters

#Tranformation applied to every channel except width and time
dataLGCLTransform <- transform(filterData,'FSC.A' = lgcl(`FSC.A`), 'SSC.A' =lgcl(`SSC.A`), 'FL1.A' = lgcl(`FL1.A`), 'FSC.H' = lgcl(`FSC.H`),'SSC.H' = lgcl(`SSC.H`),'FL1.H' = lgcl(`FL1.H`)) 




```

##Effect of time
```{r}
#The effect of time on signal (of which there shouldn't be any)
i <- 1
xyplot(FL1.A ~ Time, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(150,250), main = sampleNames(filterData)[i])
i <- 2
xyplot(FL1.A ~ Time, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(150,250), main = sampleNames(filterData)[i])
i <- 3
xyplot(FL1.A ~ Time, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(150,250), main = sampleNames(filterData)[i])
```

##Plots of FL1 versus FSC
```{r}
i <- 1
xyplot(FL1.A ~ FSC.A, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(4,8), sampleNames(filterData)[i])
i <- 2
xyplot(FL1.A ~ FSC.A, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(4,8), sampleNames(filterData)[i])
i <- 3
xyplot(FL1.A ~ FSC.A, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(4,8), sampleNames(filterData)[i])
```

##Plots of FSC versus SSC
```{r}
i <- 1
xyplot(SSC.A ~ FSC.A, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(4,8), ylim=c(4,8), sampleNames(filterData)[i])
i <- 2
xyplot(SSC.A ~ FSC.A, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(4,8), ylim=c(4,8), sampleNames(filterData)[i])
i <- 3
xyplot(SSC.A ~ FSC.A, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(4,8), ylim=c(4,8), sampleNames(filterData)[i])
```

##Extract data for further processing
```{r Extract data for further processing}

summary.stats <- as.data.frame(summary.stats)
PopProportion_zeroGFPgln <- non.gfp/non.debris
PopProportion_oneGFPgln <- gfp.one.cells/non.debris
PopProportion_twoGFPgln <- gfp.two.cells/non.debris
PopProportion_threeGFPgln <- gfp.thr.cells/non.debris
alldata <- cbind(summary.stats, PopProportion_zeroGFPgln, PopProportion_oneGFPgln, PopProportion_twoGFPgln, PopProportion_threeGFPgln, non.debris)
colnames(alldata) <- c("FSC_median","FSC_mean", "FSC_sd","FL1_median", "FL1_mean","FL1_sd","normalizedGFP_median", "normalizedGFP_mean", "normalizedGFP_sd", "PopProp_0copy", "PopProp_1copy", "PopProp_2copy", "PopProp_3plus", "Cells counted")


#Save the current dataframe
write.table(alldata, file="DataOut_112618.csv", row.names=TRUE, quote=F, sep=",")

```




