---
title: "R Notebook"
output: html_notebook
---


**Experiment overview**

Write a detailed description of your experiment here including the goal of the analysis and your interpretation of the results.   
If you still see this text it means that you have not described the experiment and whatever follows is meaningless.
###############################

> This code is designed for use with the Accuri flow cytometer, which is equiped with the following lasers and filters

* Blue laser (488 nm)
  + FL1 filter = 514/20nm   GFP
  + FL3 filter = 575/25nm   YFP

* Yellow/green laser (552 nm)
  + FL2 filter = 610/20nm   mCherry, dtomato
  + FL4 filter = 586/15nm   DsRed
  
**Requirements**  

**Output**  


#Step 1: Load relevant libraries 
```{r Libraries, eval=TRUE}
# This is a function that just makes sure you have a package, or installs it for you without prompting

requireInstall <- function(packageName,isBioconductor=F) {
  if ( !try(require(packageName,character.only=T)) ) {
    print(paste0("You don't have ",packageName," accessible, ",
      "I'm gonna install it"))
    if (isBioconductor) {
      source("http://bioconductor.org/biocLite.R")                        
      biocLite(packageName)                                                 
    } else {
      install.packages("packageName", repos = "http://cran.us.r-project.org")
    }
  }
  return(1)
}

#Load libraries
requireInstall("flowCore",isBioconductor=T)
requireInstall("flowViz",isBioconductor=T)
requireInstall("flowStats")
requireInstall("Hmisc")
requireInstall("reshape2")
requireInstall("ggplot2")
requireInstall("flowWorkspace")
requireInstall("ggcyto", isBioconductor=T)
requireInstall("gridExtra")
requireInstall("stringr")
requireInstall("devtools",isBioconductor=T)
requireInstall("ggpubr",isBioconductor=T)
devtools::install_github("guiastrennec/ggplus")
devtools::install_github("kassambara/ggpubr")
library("ggplus")
library("ggpubr")
requireInstall("tidyverse")

```
#Step 2: Read in .fcs files, an Rdata file containing the gates sample sheet(s) that contains columns with 

(Redefine Samplesheets)
* column1 = 
* column2 = 
* column3 = 
* column4 = 
* column5 = 

```{r}
#Read in all the fcs files in the directory.
path.data <-"~/Google Drive/MiniStatRun_10_2018/"

#list.folders <- c("LTEE_mCitrine_GAP1_Variants_TON", "LTEE_mCitrine_GAP1_Variants_T00", "LTEE_mCitrine_GAP1_Variants_T06", "LTEE_mCitrine_GAP1_Variants_T07", "LTEE_mCitrine_GAP1_Variants_T08.1", "LTEE_mCitrine_GAP1_Variants_T08.3", "LTEE_mCitrine_GAP1_Variants_T11.1", "LTEE_mCitrine_GAP1_Variants_T11.2", "LTEE_mCitrine_GAP1_Variants_T13.1", "LTEE_mCitrine_GAP1_Variants_T13.2", "LTEE_mCitrine_GAP1_Variants_T14", "LTEE_mCitrine_GAP1_Variants_T15", "LTEE_mCitrine_GAP1_Variants_T18")

list.folders <- c( "LTEE_mCitrine_GAP1_Variants_T00", "LTEE_mCitrine_GAP1_Variants_T06", "LTEE_mCitrine_GAP1_Variants_T07", "LTEE_mCitrine_GAP1_Variants_T08.3", "LTEE_mCitrine_GAP1_Variants_T11.2", "LTEE_mCitrine_GAP1_Variants_T13.1", "LTEE_mCitrine_GAP1_Variants_T14", "LTEE_mCitrine_GAP1_Variants_T15", "LTEE_mCitrine_GAP1_Variants_T18")


#Select folder to load, based on position in list
tp <- 1

#Set To "YES" to load all folders, tp must be set equal to 1
all <- "YES"

flowData <- read.flowSet(path = paste(path.data, list.folders[tp],"/", sep=""), pattern=".fcs", alter.names = TRUE)

sample.sheet <- read.csv(paste(path.data,"Samplesheet_",list.folders[tp],".csv", sep=""))

#Adds a sample sheet data to the pData of the flowset

sampleNames(flowData) <- paste(gsub(" ","_",sample.sheet$Strain),"_T",gsub(" ","_",sample.sheet$Timepoint),"_",sub(" ","_",sample.sheet$Well), sep="")
pData(flowData)$Strain <- gsub(" ","_",sample.sheet$Strain)
pData(flowData)$Genotype <- gsub(" ","_",sample.sheet$Genotype)
pData(flowData)$Population <- gsub(" ","_",sample.sheet$Population)
pData(flowData)$Vessel <- gsub(" ","_",sample.sheet$Vessel)
pData(flowData)$Media <- gsub(" ","_",sample.sheet$Media)
pData(flowData)$Ploidy <- gsub(" ","_",sample.sheet$Ploidy)
pData(flowData)$Well <- gsub(" ","_",sample.sheet$Well)
pData(flowData)$Experiment <- gsub(" ","_",sample.sheet$Experiment)
pData(flowData)$Timepoint <- gsub(" ","_",sample.sheet$Timepoint)
pData(flowData)$Run <- list.folders[1]

if(all == "YES") {
  for(i in 2:length(list.folders)) {
    flowData.temp <- read.flowSet(path = paste(path.data,list.folders[i],"/", sep=""), pattern=".fcs", alter.names = TRUE)

    sample.sheet <- read.csv(paste(path.data,"Samplesheet_",list.folders[i],".csv", sep=""))

    #Adds a sample sheet data to the pData of the flowset
    sampleNames(flowData.temp) <- paste(gsub(" ","_",sample.sheet$Strain),"_T",gsub(" ","_",sample.sheet$Timepoint),"_",sub(" ","_",sample.sheet$Well), sep="")
    pData(flowData.temp)$Strain <- gsub(" ","_",sample.sheet$Strain)
    pData(flowData.temp)$Genotype <- gsub(" ","_",sample.sheet$Genotype)
    pData(flowData.temp)$Population <- gsub(" ","_",sample.sheet$Population)
    pData(flowData.temp)$Vessel <- gsub(" ","_",sample.sheet$Vessel)
    pData(flowData.temp)$Media <- gsub(" ","_",sample.sheet$Media)
    pData(flowData.temp)$Ploidy <- gsub(" ","_",sample.sheet$Ploidy)
    pData(flowData.temp)$Well <- gsub(" ","_",sample.sheet$Well)
    pData(flowData.temp)$Experiment <- gsub(" ","_",sample.sheet$Experiment)
    pData(flowData.temp)$Timepoint <- gsub(" ","_",sample.sheet$Timepoint)
    pData(flowData.temp)$Run <- list.folders[i]

    flowData <- rbind2(flowData, flowData.temp)
  }
}
```

```{r flowSet summaries}
#Check how many cells were counted in each fcs file
total <- fsApply(flowData, each_col, length)[1:length(flowData)] #total counts per sample
print(total)
#Print the medians of data values for each measurement
fsApply(flowData, each_col, median)[1:length(flowData)]

#combine all flowSets into a single flowset
samples.num <- length(flowData)
print(samples.num)
```

#Step 3: apply filters to data and generate plots showing the effect on filtering
#Gating
```{r Application of Gates}
##Subset the data by applying sequential gates##

#Load Gates
load(file = "gates_LTEE_GAP1_Var.RData")

#apply doublet gate to ALL SAMPLES
subData.singlets <- Subset(subData, pg.singlets) 
fsApply(subData.singlets, each_col, length)[1:samples.num]
singlets <- fsApply(subData.singlets, each_col, length)[1:samples.num]
barplot(singlets/total, ylim=c(0,1), ylab = "Proportion singlet cells", las=2, cex.names = 0.5, names.arg=sampleNames(subData))


####GLUTAMINE GATES####
#apply debris gates
filteredData <- Subset(subData.singlets, pg.nondebris) 
fsApply(filteredData, each_col, length)[1:samples.num]
non.debris <- fsApply(filteredData, each_col, length)[1:samples.num]
barplot(non.debris/total, ylim=c(0,1), ylab = "Proportion nondebris cells", las=2, cex.names = 0.5, names.arg=str_sub(sampleNames(filteredData)))

#this gate defines nongfp cells
gfp.neg <- Subset(filteredData, gln.zero) 
fsApply(gfp.neg, each_col, length)[1:samples.num]
non.gfp <- fsApply(gfp.neg, each_col, length)[1:samples.num]
barplot(non.gfp/non.debris, ylim=c(0,1), ylab = "Proportion cells with no GFP", las=2, cex.names = 0.5, names.arg=str_sub(sampleNames(filteredData)))

#this gate defines one copy gfp cells
gfp.one <- Subset(filteredData, gln.one) 
fsApply(gfp.one, each_col, length)[1:samples.num]
gfp.one.cells <- fsApply(gfp.one, each_col, length)[1:samples.num]
barplot(gfp.one.cells/non.debris, ylim=c(0,1), ylab = "Proportion cells with 1 copy GFP", las=2, cex.names = 0.5, names.arg=str_sub(sampleNames(filteredData)))

#this gate defines two copy gfp cells
gfp.two <- Subset(filteredData, gln.two) 
fsApply(gfp.two, each_col, length)[1:samples.num]
gfp.two.cells <- fsApply(gfp.two, each_col, length)[1:samples.num]
  barplot(gfp.two.cells/non.debris, ylim=c(0,1), ylab = "Proportion cells with 2 copy GFP", las=2, cex.names = 0.5, names.arg=str_sub(sampleNames(filteredData)))

#this gate defines thre copy plus gfp cells
gfp.thr <- Subset(filteredData, gln.three) 
fsApply(gfp.thr, each_col, length)[1:samples.num]
gfp.thr.cells <- fsApply(gfp.thr, each_col, length)[1:samples.num]
barplot(gfp.thr.cells/non.debris, ylim=c(0,1), ylab = "Proportion cells with 3 copy+ GFP", las=2, cex.names = 0.5, names.arg=str_sub(sampleNames(filteredData)))
```

##Extract data from fcs files to generate statistics and boxplots
```{r Data extraction and plotting}
#Transfer data into a dataframe and produce summary stats

#Summary Statistics of Normalized Filtered Data
#Move filtered data into a dataframe
filtered.data <- data.frame(FSC.A=NA,SSC.A=NA,FL1.A=NA, FL2.A=NA, FL3.A=NA, FL4.A=NA, SAMPLE=NA, STRAIN=NA, GENOTYPE=NA, POPULATION=NA, VESSEL=NA, MEDIA=NA, PLOIDY=NA, WELL=NA, EXP=NA, TIMEPOINT=NA, RUN=NA)
stats.data <- data.frame(SAMPLE=NA,STRAIN=NA,POPULATION=NA,VESSEL=NA,TIMEPOINT=NA,COUNT=NA,FSC_MEDIAN=NA,FSC_MEAN=NA,FSC_SD=NA,FL1_MEDIAN=NA,FL1_MEAN=NA,FL1_SD=NA,NORMALIZED_GFP_MEDIAN=NA,NORMALIZED_GFP_MEAN=NA,NORMALIZED_GFP_SD=NA)

for(i in 1:length(filteredData)){
  fsc.a <- exprs(filteredData[[i,1]])
  ssc.a <- exprs(filteredData[[i,2]])
  fl1.a <- exprs(filteredData[[i,3]])
  fl2.a <- exprs(filteredData[[i,4]])
  fl3.a <- exprs(filteredData[[i,5]])
  fl4.a <- exprs(filteredData[[i,6]])
  sample <- sampleNames(filteredData)[i]
  strain <- pData(filteredData)$Strain[i]
  genotype <- pData(filteredData)$Genotype[i]
  population <- pData(filteredData)$Population[i]
  vessel <- pData(filteredData)$Vessel[i]
  media <- pData(filteredData)$Media[i]
  ploidy <- pData(filteredData)$Ploidy[i]
  well <- pData(filteredData)$Well[i]
  exp <- pData(filteredData)$Experiment[i]
  timepoint <- pData(filteredData)$Timepoint[i]
  run <- pData(filteredData)$Run[i]

  filtered.data<-rbind(filtered.data,cbind(FSC.A=fsc.a,SSC.A=ssc.a,FL1.A=fl1.a,FL2.A=fl2.a, FL3.A=fl3.a, FL4.A=fl4.a,SAMPLE=sample, STRAIN=strain, GENOTYPE=genotype, POPULATION = population, VESSEL=vessel, MEDIA=media, PLOIDY=ploidy, WELL=well, EXP=exp, TIMEPOINT=timepoint, RUN=run))
  
  stats.data<-(rbind(stats.data,cbind(SAMPLE=sample,STRAIN=strain,POPULATION=population,VESSEL=vessel,TIMEPOINT=timepoint,COUNT=length(fsc.a),FSC_MEDIAN=median(fsc.a),FSC_MEAN=mean(fsc.a),FSC_SD=sd(fsc.a),FL1_MEDIAN=median(fl1.a),FL1_MEAN=mean(fl1.a),FL1_SD=sd(fl1.a),NORMALIZED_GFP_MEDIAN=median(fl1.a/fsc.a),NORMALIZED_GFP_MEAN=mean(fl1.a/fsc.a),NORMALIZED_GFP_SD=sd(fl1.a/fsc.a))))
}  
  
 #Cleans up DataFrames
  filtered.data<-filtered.data[2:nrow(filtered.data),]
  filtered.data$FSC.A<-as.numeric(filtered.data$FSC.A)
  filtered.data$SSC.A<-as.numeric(filtered.data$SSC.A)
  filtered.data$FL1.A<-as.numeric(filtered.data$FL1.A)
  filtered.data$FL2.A<-as.numeric(filtered.data$FL2.A)
  filtered.data$FL3.A<-as.numeric(filtered.data$FL3.A)
  filtered.data$FL4.A<-as.numeric(filtered.data$FL4.A)
  filtered.data$SAMPLE<-as.factor(filtered.data$SAMPLE)
  filtered.data$STRAIN<-as.factor(filtered.data$STRAIN)
  filtered.data$GENOTYPE<-as.factor(filtered.data$GENOTYPE)
  filtered.data$POPULATION<-as.factor(filtered.data$POPULATION)
  filtered.data$VESSEL<-as.factor(filtered.data$VESSEL)
  filtered.data$MEDIA<-as.factor(filtered.data$MEDIA)
  filtered.data$PLOIDY<-as.factor(filtered.data$PLOIDY)
  filtered.data$WELL<-as.factor(filtered.data$WELL)
  filtered.data$EXP<-as.factor(filtered.data$EXP)
  filtered.data$TIMEPOINT<-as.factor(filtered.data$TIMEPOINT)
  filtered.data$RUN<-as.factor(filtered.data$RUN)
  
  
  stats.data<-stats.data[2:nrow(stats.data),]
  stats.data$SAMPLE<-as.factor(stats.data$SAMPLE)
  stats.data$STRAIN<-as.factor(stats.data$STRAIN)
  stats.data$POPULATION<-as.factor(stats.data$POPULATION)
  stats.data$VESSEL<-as.factor(stats.data$VESSEL)
  stats.data$TIMEPOINT<-as.factor(stats.data$TIMEPOINT)
  stats.data$COUNT<-as.numeric(stats.data$COUNT)
  stats.data$FSC_MEDIAN<-as.numeric(stats.data$FSC_MEDIAN)
  stats.data$FSC_MEAN<-as.numeric(stats.data$FSC_MEAN)    
  stats.data$FSC_SD<-as.numeric(stats.data$FSC_SD)
  stats.data$FL1_MEDIAN<-as.numeric(stats.data$FL1_MEDIAN)
  stats.data$FL1_MEAN<-as.numeric(stats.data$FL1_MEAN)    
  stats.data$FL1_SD<-as.numeric(stats.data$FL1_SD)
  stats.data$NORMALIZED_GFP_MEDIAN<-as.numeric(stats.data$NORMALIZED_GFP_MEDIAN)
  stats.data$NORMALIZED_GFP_MEAN<-as.numeric(stats.data$NORMALIZED_GFP_MEAN)    
  stats.data$NORMALIZED_GFP_SD<-as.numeric(stats.data$NORMALIZED_GFP_SD)
  
```

##Extract data for further processing
```{r Extract data for further processing}

PopProportion_zeroGFP <- non.gfp/non.debris
PopProportion_oneGFP <- gfp.one.cells/non.debris
PopProportion_twoGFP <- gfp.two.cells/non.debris
PopProportion_threeGFP <- gfp.thr.cells/non.debris
all.data <- cbind(stats.data, PopProportion_zeroGFP, PopProportion_oneGFP, PopProportion_twoGFP, PopProportion_threeGFP, non.debris)
colnames(all.data) <- c(colnames(stats.data), "PopProp_0copy", "PopProp_1copy", "PopProp_2copy", "PopProp_3plus", "Cells counted")


#Save the current dataframe
write.table(all.data, file="DataOut_11302018.csv", row.names=TRUE, quote=F, sep=",")
```

