---
title: "Gresham Lab Flow Core Guide"
author: '`r Sys.info()[7]`'
date: '`r Sys.Date()`'
output:
  html_document:
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
---


**Experiment overview**

Write a detailed description of your experiment here including the goal of the analysis and your interpretation of the results.   
If you still see this text it means that you have not described the experiment and whatever follows is meaningless.
###############################

> This code is designed for use with the Accuri flow cytometer, which is equiped with the following lasers and filters

* Blue laser (488 nm)
  + FL1 filter = 514/20nm   GFP
  + FL3 filter = 575/25nm   YFP

* Yellow/green laser (552 nm)
  + FL2 filter = 610/20nm   mCherry, dtomato
  + FL4 filter = 586/15nm   DsRed
  
  

**Requirements**  
In order to run this code you need:  
  + to predefine your gates using the **gating.R** script  
  + the **gates.Rdata** workspace, which contains the gates used in this script  
  + the path of the directory(ies), given the variable names **dir1**, **dir2**... that contain .fcs files named A01.fcs, A02.fcs, A03.fcs...  
  + a tab delimited sample sheet in each directory with the following rows: <Well>	<Strain>	<Genotype>	<Ploidy>	<Media>	<Experiment>  
  + the variable names are changed in chunk 2 named "Variable Names"    




**Output**  
This script generates a summary of results followed by quality control plots.  



#Step 1: Load relevant libraries 
```{r Libraries, eval=TRUE}
# This is a function that just makes sure you have a package, or installs it for you without prompting

requireInstall <- function(packageName,isBioconductor=F) {
  if ( !try(require(packageName,character.only=T)) ) {
    print(paste0("You don't have ",packageName," accessible, ",
      "I'm gonna install it"))
    if (isBioconductor) {
      source("http://bioconductor.org/biocLite.R")                        
      biocLite(packageName)                                                 
    } else {
      install.packages("packageName", repos = "http://cran.us.r-project.org")
    }
  }
  return(1)
}

#Load libraries
requireInstall("flowCore",isBioconductor=T)
requireInstall("flowViz",isBioconductor=T)
requireInstall("flowStats")
requireInstall("Hmisc")
requireInstall("reshape2")
requireInstall("ggplot2")
requireInstall("flowWorkspace")
requireInstall("ggcyto", isBioconductor=T)
requireInstall("gridExtra")


```

#Step 2: Read in .fcs files, an Rdata file containing the gates sample sheet(s) that contains four columns with 
* column1 = Well
* column2 = Strain
* column3 = Staining
* column4 = Media
* column5 = Userdefined

```{r Variable definitions}
#Read in all data for analysis. Data should be in individual directories that contain .fcs files and a corresponding sample sheet with a generic format. FCS file names should be unaltered e.g AO1.fcs, A02.fcs, ...H12.fcs 
#An abitrary number of directories can be used named dir1, dir2, dir3...with a corresponding flowData.1, flowData.2, flowData.3...and sample.sheet.1, sample.sheet.2, sample.sheet.3...

#load the Rdata file containing the gates
load("gates_05252017_MIDTPT.Rdata") 

#Read in all the fcs files in the directory, with alter.names changing "-" to "."  
flowData <- read.flowSet(path = ".", pattern=".fcs", alter.names=TRUE)

#Read in the sample sheet that should be in each directory that contains the .fcs files.  
sample.sheet <- read.delim(paste("SampleSheet.txt", sep="/"))

#Change names of samples to those specified in the sample sheets
sampleNames(flowData) <- paste(sample.sheet[,1], sample.sheet[,2], sample.sheet[,3], sample.sheet[,4], sample.sheet[,5], sep=" ")
```


```{r flowSet summaries}
#Check how many cells were counted in each fcs file
fsApply(flowData, each_col, length)[1:32]

total <- fsApply(flowData, each_col, length)[1:32]  #total counts per sample

#Print the medians of data values for each measurement
fsApply(flowData, each_col, median)

#combine all flowSets into a single flowset
samples.num <- length(flowData)

```

#Step 3: apply filters to data and generate plots showing the effect on filtering
```{r Application of Gates}
##Subset the data by applying sequential gates##

#apply doublet gate to ALL SAMPLES
flowData.singlets <- Subset(flowData, pg.singlets) 
fsApply(flowData.singlets, each_col, length)[1:samples.num]
singlets <- fsApply(flowData.singlets, each_col, length)[1:samples.num]
barplot(singlets/total, ylim=c(0,1), ylab = "Proportion singlet cells", las=2, cex.names = 0.5, names.arg=sampleNames(flowData))

##Subset data by media type
flowGlu <- flowSet(flowData[[1]], flowData[[2]], flowData[[3]], flowData[[9]], flowData[[10]], flowData[[11]], flowData[[17]], flowData[[18]], flowData[[25]], flowData[[26]])
total.glu <- fsApply(flowGlu, each_col, length)[1:10]
samples.num.glu <- length(flowGlu)
sample.sheet.glu <- read.delim(paste("SampleSheetglu.txt", sep="/"))
sampleNames(flowGlu) <- paste(sample.sheet.glu[,1], sample.sheet.glu[,2], sample.sheet.glu[,3], sample.sheet.glu[,4], sample.sheet.glu[,5], sep=" ")

flowGlu.singlets <- Subset(flowGlu, pg.singletsGLU) 
fsApply(flowGlu.singlets, each_col, length)[1:samples.num.glu]
GLUsinglets <- fsApply(flowGlu.singlets, each_col, length)[1:samples.num.glu]
barplot(GLUsinglets/total.glu, ylim=c(0,1), ylab = "Proportion GLUCOSE singlet cells", las=2, cex.names = 0.5, names.arg=sampleNames(flowGlu))

flowUr <- flowSet(flowData.singlets[[4]], flowData.singlets[[5]], flowData.singlets[[6]], flowData.singlets[[12]], flowData.singlets[[13]], flowData.singlets[[19]], flowData.singlets[[20]], flowData.singlets[[21]], flowData.singlets[[27]], flowData.singlets[[28]], flowData.singlets[[29]])
total.ur <- fsApply(flowUr, each_col, length)[1:11]
samples.num.ur <-length(flowUr)
sample.sheet.ur <- read.delim(paste("SampleSheetur.txt", sep="/"))
sampleNames(flowUr) <- paste(sample.sheet.ur[,1], sample.sheet.ur[,2], sample.sheet.ur[,3], sample.sheet.ur[,4], sample.sheet.ur[,5], sep=" ")

flowGln <- flowSet(flowData.singlets[[7]], flowData.singlets[[8]], flowData.singlets[[14]], flowData.singlets[[15]], flowData.singlets[[16]], flowData.singlets[[22]], flowData.singlets[[23]], flowData.singlets[[24]], flowData.singlets[[30]], flowData.singlets[[31]], flowData.singlets[[32]])
total.gln <- fsApply(flowGln, each_col, length)[1:11]
samples.num.gln <- length(flowGln)
sample.sheet.gln <- read.delim(paste("SampleSheetgln.txt", sep="/"))
sampleNames(flowGln) <- paste(sample.sheet.gln[,1], sample.sheet.gln[,2], sample.sheet.gln[,3], sample.sheet.gln[,4], sample.sheet.gln[,5], sep=" ")


###GLUCOSE GATES###
#apply debris gates
filteredGluData <- Subset(flowGlu.singlets, pg.nondebris.glu) 
fsApply(filteredGluData, each_col, length)[1:samples.num.glu]
non.debris.glu <- fsApply(filteredGluData, each_col, length)[1:samples.num.glu]
barplot(non.debris.glu/total.glu, ylim=c(0,1), ylab = "Proportion nondebris cells", las=2, cex.names = 0.5, names.arg=sampleNames(flowGlu))

#this gate defines nongfp cells
gfp.neg.glu <- Subset(filteredGluData, pg.zcht) 
fsApply(gfp.neg.glu, each_col, length)[1:samples.num.glu]
non.gfp.glu <- fsApply(gfp.neg.glu, each_col, length)[1:samples.num.glu]
barplot(non.gfp.glu/non.debris.glu, ylim=c(0,1), ylab = "Proportion cells with no GFP", las=2, cex.names = 0.5, names.arg=sampleNames(flowGlu))

#this gate defines one copy gfp cells
gfp.one.glu <- Subset(filteredGluData, pg.ocht) 
fsApply(gfp.one.glu, each_col, length)[1:samples.num.glu]
gfp.one.cells.glu <- fsApply(gfp.one.glu, each_col, length)[1:samples.num.glu]
barplot(gfp.one.cells.glu/non.debris.glu, ylim=c(0,1), ylab = "Proportion cells with 1 copy GFP", las=2, cex.names = 0.5, names.arg=sampleNames(flowGlu))

#this gate defines two copy gfp cells
gfp.two.glu <- Subset(filteredGluData, pg.twcht) 
fsApply(gfp.two.glu, each_col, length)[1:samples.num.glu]
gfp.two.cells.glu <- fsApply(gfp.two.glu, each_col, length)[1:samples.num.glu]
barplot(gfp.two.cells.glu/non.debris.glu, ylim=c(0,1), ylab = "Proportion cells with 2 copy GFP", las=2, cex.names = 0.5, names.arg=sampleNames(flowGlu))

#this gate defines thre copy plus gfp cells
gfp.thr.glu <- Subset(filteredGluData, pg.thrcht) 
fsApply(gfp.thr.glu, each_col, length)[1:samples.num.glu]
gfp.thr.cells.glu <- fsApply(gfp.thr.glu, each_col, length)[1:samples.num.glu]
barplot(gfp.thr.cells.glu/non.debris.glu, ylim=c(0,1), ylab = "Proportion cells with 3 copy+ GFP", las=2, cex.names = 0.5, names.arg=sampleNames(flowGlu))

###UREA GATES###
#apply debris gates
filteredUrData <- Subset(flowUr, pg.nondebris.ur) 
fsApply(filteredUrData, each_col, length)[1:samples.num.ur]
non.debris.ur <- fsApply(filteredUrData, each_col, length)[1:samples.num.ur]
barplot(non.debris.ur/total.ur, ylim=c(0,1), ylab = "Proportion nondebris cells", las=2, cex.names = 0.5, names.arg=sampleNames(flowUr))

#this gate defines nongfp cells
gfp.neg.ur <- Subset(filteredUrData, ur.zcht) 
fsApply(gfp.neg.ur, each_col, length)[1:samples.num.ur]
non.gfp.ur <- fsApply(gfp.neg.ur, each_col, length)[1:samples.num.ur]
barplot(non.gfp.ur/non.debris.ur, ylim=c(0,1), ylab = "Proportion cells with no GFP", las=2, cex.names = 0.5, names.arg=sampleNames(flowUr))

#this gate defines one copy gfp cells
gfp.one.ur <- Subset(filteredUrData, ur.ocht) 
fsApply(gfp.one.ur, each_col, length)[1:samples.num.ur]
gfp.one.cells.ur <- fsApply(gfp.one.ur, each_col, length)[1:samples.num.ur]
barplot(gfp.one.cells.ur/non.debris.ur, ylim=c(0,1), ylab = "Proportion cells with 1 copy GFP", las=2, cex.names = 0.5, names.arg=sampleNames(flowUr))

#this gate defines two copy gfp cells
gfp.two.ur <- Subset(filteredUrData, ur.twcht) 
fsApply(gfp.two.ur, each_col, length)[1:samples.num.ur]
gfp.two.cells.ur <- fsApply(gfp.two.ur, each_col, length)[1:samples.num.ur]
barplot(gfp.two.cells.ur/non.debris.ur, ylim=c(0,1), ylab = "Proportion cells with 2 copy GFP", las=2, cex.names = 0.5, names.arg=sampleNames(flowUr))

#this gate defines thre copy plus gfp cells
gfp.thr.ur <- Subset(filteredUrData, ur.thrcht) 
fsApply(gfp.thr.ur, each_col, length)[1:samples.num.ur]
gfp.thr.cells.ur <- fsApply(gfp.thr.ur, each_col, length)[1:samples.num.ur]
barplot(gfp.thr.cells.ur/non.debris.ur, ylim=c(0,1), ylab = "Proportion cells with 3 copy+ GFP", las=2, cex.names = 0.5, names.arg=sampleNames(flowUr))

###GLUTAMINE GATES###
#apply debris gates
filteredGlnData <- Subset(flowGln, pg.nondebris.gln) 
fsApply(filteredGlnData, each_col, length)[1:samples.num.gln]
non.debris.gln <- fsApply(filteredGlnData, each_col, length)[1:samples.num.gln]
barplot(non.debris.gln/total.gln, ylim=c(0,1), ylab = "Proportion nondebris cells", las=2, cex.names = 0.5, names.arg=sampleNames(flowGln))

#this gate defines nongfp cells
gfp.neg.gln <- Subset(filteredGlnData, gln.zcht) 
fsApply(gfp.neg.gln, each_col, length)[1:samples.num.gln]
non.gfp.gln <- fsApply(gfp.neg.gln, each_col, length)[1:samples.num.gln]
barplot(non.gfp.gln/non.debris.gln, ylim=c(0,1), ylab = "Proportion cells with no GFP", las=2, cex.names = 0.5, names.arg=sampleNames(flowGln))

#this gate defines one copy gfp cells
gfp.one.gln <- Subset(filteredGlnData, gln.ocht) 
fsApply(gfp.one.gln, each_col, length)[1:samples.num.gln]
gfp.one.cells.gln <- fsApply(gfp.one.gln, each_col, length)[1:samples.num.gln]
barplot(gfp.one.cells.gln/non.debris.gln, ylim=c(0,1), ylab = "Proportion cells with 1 copy GFP", las=2, cex.names = 0.5, names.arg=sampleNames(flowGln))

#this gate defines two copy gfp cells
gfp.two.gln <- Subset(filteredGlnData, gln.twcht) 
fsApply(gfp.two.gln, each_col, length)[1:samples.num.gln]
gfp.two.cells.gln <- fsApply(gfp.two.gln, each_col, length)[1:samples.num.gln]
barplot(gfp.two.cells.gln/non.debris.gln, ylim=c(0,1), ylab = "Proportion cells with 2 copy GFP", las=2, cex.names = 0.5, names.arg=sampleNames(flowGln))

#this gate defines thre copy plus gfp cells
gfp.thr.gln <- Subset(filteredGlnData, gln.thrcht) 
fsApply(gfp.thr.gln, each_col, length)[1:samples.num.gln]
gfp.thr.cells.gln <- fsApply(gfp.thr.gln, each_col, length)[1:samples.num.gln]
barplot(gfp.thr.cells.gln/non.debris.gln, ylim=c(0,1), ylab = "Proportion cells with 3 copy+ GFP", las=2, cex.names = 0.5, names.arg=sampleNames(flowGln))
```

#Step 4: Data analysis

##diagnostic values can be defined for plotting purposes
```{r Definition of diagnostic values}
#define critical values that can superimposed on plots for easy visual comparison

gfp.bg <- 3.9  #a background value for GFP
gfp.wt <- 5.9 #a value for wildtype GFP expression
haploid.fsc <- 6e5 #an empirical value for forward scatter for haploids
diploid.fsc <- 7e5 #an empirical value for forward scatter for diploids
gfp.norm <- 0.935 #an empricial value for gfp expression normalized by forward scatter

```

##Extract data from fcs files to generate statistics and boxplots
```{r Data extraction and plotting}
###GLUCOSE###
#record summary statistics for each sample in a matrix named summary.stats 
summary.stats.glu <- matrix(data = NA, nrow = length(filteredGluData), ncol = 9, dimnames = list(sampleNames(filteredGluData),c("FSC_median","FSC_mean", "FSC_sd","FL1_median", "FL1_mean","FL1_sd","normalizedGFP_median", "normalizedGFP_mean", "normalizedGFP_sd")))

#use the sample containing the minimum number of points after filtering for doublets and debris to define the number of data points retained for all samples
sample.size.glu <- min(fsApply(filteredGluData, each_col, length))  

print(sample.size.glu)

comparison.FSC.glu <- matrix(data = NA, nrow = sample.size.glu, ncol = length(filteredGluData), byrow = FALSE,dimnames = NULL)
comparison.FL1.glu <- matrix(data = NA, nrow = sample.size.glu, ncol = length(filteredGluData), byrow = FALSE,dimnames = NULL)
comparison.FL1NormFsc.glu <- matrix(data = NA, nrow = sample.size.glu, ncol = length(filteredGluData), byrow = FALSE,dimnames = NULL)

#for each sample plot a histogram of the normalized data, raw FSC and raw GFP per row
par(mfrow=c(1,2), mar=c(5.1,2.1,2.1,2.1), oma=c(1.5,2,1,1))

#extract data from flowFrames to plot histograms of values and record summary statistics
for (i in 1:length(filteredGluData)){
 
  temp <- exprs(filteredGluData[[i]]) #exprs() extracts a matrix of the values from the flowframe
 

  ##########################################
  #record summary statistics for the sample#
  ##########################################
  
  #FSC
  summary.stats.glu[i,1] <- median(temp[,1]) 
  summary.stats.glu[i,2] <-mean(temp[,1])  
  summary.stats.glu[i,3] <- sd(temp[,1])
  #FL1
  summary.stats.glu[i,4] <- median(temp[,3])
  summary.stats.glu[i,5] <-mean(temp[,3])  
  summary.stats.glu[i,6] <- sd(temp[,3])
  #FL1 (GFP) divided by FSC
  summary.stats.glu[i,7] <- median(temp[,3]/temp[,1])
  summary.stats.glu[i,8] <-mean(temp[,3]/temp[,1])  
  summary.stats.glu[i,9] <- sd(temp[,3]/temp[,1])
  
  ##############################################
  #plot histograms of the channels of interest##
  ##############################################

  ###############
  #Green channel#
  ###############
  
  #FL1 (GFP)
  hist(log10(temp[,3]), br=1000, xlab = "log10(FL1)", main = "FL1") 
  abline(v=gfp.bg, col="yellow", lty=2, lwd=2)
  abline(v=gfp.wt, col="green", lty=2, lwd=2) 
  legend("topleft",  legend=paste("median FL1 = ",round(median(temp[,3]), digits=4),sep=""))

  #GFP divided by FSC
  hist(temp[,3]/temp[,1], br=500, xlab = "FL1/FSC", main = "FL1/FSC") 
  abline(v=gfp.norm, col="green", lty=2, lwd=2 )
  legend("topleft",  legend=paste("median GFP / FSC=",round(median(temp[,3]/temp[,1]), digits=4),sep=""))
  
  mtext(sampleNames(flowGlu[i]), outer = TRUE, cex = 1.0)
  
  ###############
  #Other#########
  ###############

    #FSC
  hist(log10(temp[,1]), br=500, xlab = "log10(FSC)", main = "FSC", xlim=c(4,8)) 
  abline(v=haploid.fsc, col="blue", lty=2, lwd=2)
  abline(v=diploid.fsc, col="grey", lty=2, lwd=2)
  legend("topleft",  legend=paste("median FSC=",round(median(temp[,1]), digits=4),sep=""))
  
  mtext(sampleNames(flowGlu[i]), outer = TRUE, cex = 1.0)

print("-------------------------------------------------------")
print("-----------------------------------")
print("----------------------")

  ############################################################
  #keep the data set for generating boxplots comparing values#
  ############################################################
  
  #Note that the amount of data kept for each sample is defined by the lowest count among all the samples.
  comparison.FSC.glu[1:sample.size.glu,i] <- temp[1:sample.size.glu,1] #FSC
  comparison.FL1.glu[1:sample.size.glu,i] <- temp[1:sample.size.glu,3] #FL1 (GFP)
  comparison.FL1NormFsc.glu[1:sample.size.glu,i] <- temp[1:sample.size.glu,3]/temp[1:sample.size.glu,1] #GFP/FSC
}

###UREA###
#record summary statistics for each sample in a matrix named summary.stats 
summary.stats.ur <- matrix(data = NA, nrow = length(filteredUrData), ncol = 9, dimnames = list(sampleNames(filteredUrData),c("FSC_median","FSC_mean", "FSC_sd","FL1_median", "FL1_mean","FL1_sd","normalizedGFP_median", "normalizedGFP_mean", "normalizedGFP_sd")))

#use the sample containing the minimum number of points after filtering for doublets and debris to define the number of data points retained for all samples
sample.size.ur <- min(fsApply(filteredUrData, each_col, length))  

print(sample.size.ur)

comparison.FSC.ur <- matrix(data = NA, nrow = sample.size.ur, ncol = length(filteredUrData), byrow = FALSE,dimnames = NULL)
comparison.FL1.ur <- matrix(data = NA, nrow = sample.size.ur, ncol = length(filteredUrData), byrow = FALSE,dimnames = NULL)
comparison.FL1NormFsc.ur <- matrix(data = NA, nrow = sample.size.ur, ncol = length(filteredUrData), byrow = FALSE,dimnames = NULL)

#for each sample plot a histogram of the normalized data, raw FSC and raw GFP per row
par(mfrow=c(1,2), mar=c(5.1,2.1,2.1,2.1), oma=c(1.5,2,1,1))

#extract data from flowFrames to plot histograms of values and record summary statistics
for (i in 1:length(filteredUrData)){
 
  temp <- exprs(filteredUrData[[i]]) #exprs() extracts a matrix of the values from the flowframe
 

  ##########################################
  #record summary statistics for the sample#
  ##########################################
  
  #FSC
  summary.stats.ur[i,1] <- median(temp[,1]) 
  summary.stats.ur[i,2] <-mean(temp[,1])  
  summary.stats.ur[i,3] <- sd(temp[,1])
  #FL1
  summary.stats.ur[i,4] <- median(temp[,3])
  summary.stats.ur[i,5] <-mean(temp[,3])  
  summary.stats.ur[i,6] <- sd(temp[,3])
  #FL1 (GFP) divided by FSC
  summary.stats.ur[i,7] <- median(temp[,3]/temp[,1])
  summary.stats.ur[i,8] <-mean(temp[,3]/temp[,1])  
  summary.stats.ur[i,9] <- sd(temp[,3]/temp[,1])
  
  ##############################################
  #plot histograms of the channels of interest##
  ##############################################

  ###############
  #Green channel#
  ###############
  
  #FL1 (GFP)
  hist(log10(temp[,3]), br=1000, xlab = "log10(FL1)", main = "FL1") 
  abline(v=gfp.bg, col="yellow", lty=2, lwd=2)
  abline(v=gfp.wt, col="green", lty=2, lwd=2) 
  legend("topleft",  legend=paste("median FL1 = ",round(median(temp[,3]), digits=4),sep=""))

  #GFP divided by FSC
  hist(temp[,3]/temp[,1], br=500, xlab = "FL1/FSC", main = "FL1/FSC") 
  abline(v=gfp.norm, col="green", lty=2, lwd=2 )
  legend("topleft",  legend=paste("median GFP / FSC=",round(median(temp[,3]/temp[,1]), digits=4),sep=""))
  
  mtext(sampleNames(flowUr[i]), outer = TRUE, cex = 1.0)
  
  ###############
  #Other#########
  ###############

    #FSC
  hist(log10(temp[,1]), br=500, xlab = "log10(FSC)", main = "FSC", xlim=c(4,8)) 
  abline(v=haploid.fsc, col="blue", lty=2, lwd=2)
  abline(v=diploid.fsc, col="grey", lty=2, lwd=2)
  legend("topleft",  legend=paste("median FSC=",round(median(temp[,1]), digits=4),sep=""))
  
  mtext(sampleNames(flowUr[i]), outer = TRUE, cex = 1.0)

print("-------------------------------------------------------")
print("-----------------------------------")
print("----------------------")

  ############################################################
  #keep the data set for generating boxplots comparing values#
  ############################################################
  
  #Note that the amount of data kept for each sample is defined by the lowest count among all the samples.
  comparison.FSC.ur[1:sample.size.ur,i] <- temp[1:sample.size.ur,1] #FSC
  comparison.FL1.ur[1:sample.size.ur,i] <- temp[1:sample.size.ur,3] #FL1 (GFP)
  comparison.FL1NormFsc.ur[1:sample.size.ur,i] <- temp[1:sample.size.ur,3]/temp[1:sample.size.ur,1] #GFP/FSC
}

par(mfrow=c(1,1)) #change number of plots per row back to standard

###GLUTAMINE###
#record summary statistics for each sample in a matrix named summary.stats 
summary.stats.gln <- matrix(data = NA, nrow = length(filteredGlnData), ncol = 9, dimnames = list(sampleNames(filteredGlnData),c("FSC_median","FSC_mean", "FSC_sd","FL1_median", "FL1_mean","FL1_sd","normalizedGFP_median", "normalizedGFP_mean", "normalizedGFP_sd")))

#use the sample containing the minimum number of points after filtering for doublets and debris to define the number of data points retained for all samples
sample.size.gln <- min(fsApply(filteredGlnData, each_col, length))  

print(sample.size.gln)

comparison.FSC.gln <- matrix(data = NA, nrow = sample.size.gln, ncol = length(filteredGlnData), byrow = FALSE,dimnames = NULL)
comparison.FL1.gln <- matrix(data = NA, nrow = sample.size.gln, ncol = length(filteredGlnData), byrow = FALSE,dimnames = NULL)
comparison.FL1NormFsc.gln <- matrix(data = NA, nrow = sample.size.gln, ncol = length(filteredGlnData), byrow = FALSE,dimnames = NULL)

#for each sample plot a histogram of the normalized data, raw FSC and raw GFP per row
par(mfrow=c(1,2), mar=c(5.1,2.1,2.1,2.1), oma=c(1.5,2,1,1))

#extract data from flowFrames to plot histograms of values and record summary statistics
for (i in 1:length(filteredGlnData)){
 
  temp <- exprs(filteredGlnData[[i]]) #exprs() extracts a matrix of the values from the flowframe
 

  ##########################################
  #record summary statistics for the sample#
  ##########################################
  
  #FSC
  summary.stats.gln[i,1] <- median(temp[,1]) 
  summary.stats.gln[i,2] <-mean(temp[,1])  
  summary.stats.gln[i,3] <- sd(temp[,1])
  #FL1
  summary.stats.gln[i,4] <- median(temp[,3])
  summary.stats.gln[i,5] <-mean(temp[,3])  
  summary.stats.gln[i,6] <- sd(temp[,3])
  #FL1 (GFP) divided by FSC
  summary.stats.gln[i,7] <- median(temp[,3]/temp[,1])
  summary.stats.gln[i,8] <-mean(temp[,3]/temp[,1])  
  summary.stats.gln[i,9] <- sd(temp[,3]/temp[,1])
  
  ##############################################
  #plot histograms of the channels of interest##
  ##############################################

  ###############
  #Green channel#
  ###############
  
  #FL1 (GFP)
  hist(log10(temp[,3]), br=1000, xlab = "log10(FL1)", main = "FL1") 
  abline(v=gfp.bg, col="yellow", lty=2, lwd=2)
  abline(v=gfp.wt, col="green", lty=2, lwd=2) 
  legend("topleft",  legend=paste("median FL1 = ",round(median(temp[,3]), digits=4),sep=""))

  #GFP divided by FSC
  hist(temp[,3]/temp[,1], br=500, xlab = "FL1/FSC", main = "FL1/FSC") 
  abline(v=gfp.norm, col="green", lty=2, lwd=2 )
  legend("topleft",  legend=paste("median GFP / FSC=",round(median(temp[,3]/temp[,1]), digits=4),sep=""))
  
  mtext(sampleNames(flowGln[i]), outer = TRUE, cex = 1.0)
  
  ###############
  #Other#########
  ###############

    #FSC
  hist(log10(temp[,1]), br=500, xlab = "log10(FSC)", main = "FSC", xlim=c(4,8)) 
  abline(v=haploid.fsc, col="blue", lty=2, lwd=2)
  abline(v=diploid.fsc, col="grey", lty=2, lwd=2)
  legend("topleft",  legend=paste("median FSC=",round(median(temp[,1]), digits=4),sep=""))
  
  mtext(sampleNames(flowGln[i]), outer = TRUE, cex = 1.0)

print("-------------------------------------------------------")
print("-----------------------------------")
print("----------------------")

  ############################################################
  #keep the data set for generating boxplots comparing values#
  ############################################################
  
  #Note that the amount of data kept for each sample is defined by the lowest count among all the samples.
  comparison.FSC.gln[1:sample.size.gln,i] <- temp[1:sample.size.gln,1] #FSC
  comparison.FL1.gln[1:sample.size.gln,i] <- temp[1:sample.size.gln,3] #FL1 (GFP)
  comparison.FL1NormFsc.gln[1:sample.size.gln,i] <- temp[1:sample.size.gln,3]/temp[1:sample.size.gln,1] #GFP/FSC
}

```


##Overview of data distributions
```{r Overall data distributions}
par(mar=c(8.1,4.1,4.1,2.1)) #create more space at lower margin

###GLUCOSE###
boxplot(comparison.FSC.glu, names=sampleNames(flowGlu), notch = TRUE, col = "gray", ylab="FSC", cex.axis=0.5,las=2, outline=F)
abline(h=haploid.fsc, lty=2, col=2)
abline(h=diploid.fsc, lty=2, col=3)

boxplot(comparison.FL1.glu, names=sampleNames(flowGlu), notch = TRUE, col = "lightgreen", ylab="FL1", cex.axis=0.5,las=2, outline=F)
abline(h=gfp.bg ,lty=2, lwd=3, col="yellow")
abline(h=gfp.wt, lty = 2, lwd=3, col="green")

boxplot(comparison.FL1NormFsc.glu, names=sampleNames(flowGlu), notch = TRUE, col = "green", ylab="FL1/FSC", cex.axis=0.5,las=2, outline=F)
abline(h=gfp.norm, lty=2, lwd=3, col="blue")

###UREA###
boxplot(comparison.FSC.ur, names=sampleNames(flowUr), notch = TRUE, col = "gray", ylab="FSC", cex.axis=0.5,las=2, outline=F)
abline(h=haploid.fsc, lty=2, col=2)
abline(h=diploid.fsc, lty=2, col=3)

boxplot(comparison.FL1.ur, names=sampleNames(flowUr), notch = TRUE, col = "lightgreen", ylab="FL1", cex.axis=0.5,las=2, outline=F)
abline(h=gfp.bg ,lty=2, lwd=3, col="yellow")
abline(h=gfp.wt, lty = 2, lwd=3, col="green")

boxplot(comparison.FL1NormFsc.ur, names=sampleNames(flowUr), notch = TRUE, col = "green", ylab="FL1/FSC", cex.axis=0.5,las=2, outline=F)
abline(h=gfp.norm, lty=2, lwd=3, col="blue")

###GLUTAMINE###
boxplot(comparison.FSC.gln, names=sampleNames(flowGln), notch = TRUE, col = "gray", ylab="FSC", cex.axis=0.5,las=2, outline=F)
abline(h=haploid.fsc, lty=2, col=2)
abline(h=diploid.fsc, lty=2, col=3)

boxplot(comparison.FL1.gln, names=sampleNames(flowGln), notch = TRUE, col = "lightgreen", ylab="FL1", cex.axis=0.5,las=2, outline=F)
abline(h=gfp.bg ,lty=2, lwd=3, col="yellow")
abline(h=gfp.wt, lty = 2, lwd=3, col="green")

boxplot(comparison.FL1NormFsc.gln, names=sampleNames(flowGln), notch = TRUE, col = "green", ylab="FL1/FSC", cex.axis=0.5,las=2, outline=F)
abline(h=gfp.norm, lty=2, lwd=3, col="blue")

par(mar=c(5.1,4.1,4.1,2.1)) #reset margins to default

#generate summary tables containing all the recorded statistics
print(summary.stats.glu)
summary.stats.glu <- as.data.frame(summary.stats.glu)
print(summary.stats.ur)
summary.stats.ur <- as.data.frame(summary.stats.ur)
print(summary.stats.gln)
summary.stats.gln <- as.data.frame(summary.stats.gln)

```


##Quantitation of relative FL1 signal
```{r }
###GLUCOSE###
baseline.FL1.glu <- summary.stats.glu$FL1_median[3] #MUST REPLACE WITH YOUR 1 COPY CONTROL!!
barplot(summary.stats.glu$FL1_median/baseline.FL1.glu, ylim = c(0,2), ylab="Relative FL1 median expression", las=2, cex.names = 0.5, names.arg=sampleNames(flowGlu))

###UREA###
baseline.FL1.ur <- summary.stats.ur$FL1_median[11] #MUST REPLACE WITH YOUR 1 COPY CONTROL!!
barplot(summary.stats.ur$FL1_median/baseline.FL1.ur, ylim = c(0,2), ylab="Relative FL1 median expression", las=2, cex.names = 0.5, names.arg=sampleNames(flowUr))

###GLUTAMINE###
baseline.FL1.gln <- summary.stats.gln$FL1_median[11] #MUST REPLACE WITH YOUR 1 COPY CONTROL!!
barplot(summary.stats.gln$FL1_median/baseline.FL1.gln, ylim = c(0,2), ylab="Relative FL1 median expression", las=2, cex.names = 0.5, names.arg=sampleNames(flowGln))

```

##Quantitation of forward scatter
```{r}

###GLUCOSE###
baseline.FSC.glu <- summary.stats.glu$FSC_median[3]
barplot(summary.stats.glu$FSC_median/baseline.FSC.glu, ylab="Relative median FSC", las=2, cex.names = 0.5, names.arg=sampleNames(flowGlu))

###UREA###
baseline.FSC.ur <- summary.stats.ur$FSC_median[11]
barplot(summary.stats.ur$FSC_median/baseline.FSC.ur, ylab="Relative median FSC", las=2, cex.names = 0.5, names.arg=sampleNames(flowUr))

###GLUTAMINE###
baseline.FSC.gln <- summary.stats.gln$FSC_median[11]
barplot(summary.stats.gln$FSC_median/baseline.FSC.gln, ylab="Relative median FSC", las=2, cex.names = 0.5, names.arg=sampleNames(flowGln))

```

##Population composition
```{r}
###GLUCOSE###
pop.composition.glu <- rbind(non.gfp.glu/non.debris.glu,gfp.one.cells.glu/non.debris.glu,gfp.two.cells.glu/non.debris.glu, gfp.thr.cells.glu/non.debris.glu)
barplot(pop.composition.glu, ylab="Proportion of population", legend=c("No GFP", "One copy GFP", "Two copy GFP", "Three+ copy GFP"),las=2, ylim=c(0,1), col=c("black", "grey", "light green", "dark green"), cex.names = 0.5,names.arg=sampleNames(flowGlu))

###UREA###
pop.composition.ur <- rbind(non.gfp.ur/non.debris.ur,gfp.one.cells.ur/non.debris.ur,gfp.two.cells.ur/non.debris.ur, gfp.thr.cells.ur/non.debris.ur)
barplot(pop.composition.ur, ylab="Proportion of population", legend=c("No GFP", "One copy GFP", "Two copy GFP", "Three+ copy GFP"),las=2, ylim=c(0,1), col=c("black", "grey", "light green", "dark green"), cex.names = 0.5,names.arg=sampleNames(flowUr))

###GLUTAMINE###
pop.composition.gln <- rbind(non.gfp.gln/non.debris.gln,gfp.one.cells.gln/non.debris.gln,gfp.two.cells.gln/non.debris.gln, gfp.thr.cells.gln/non.debris.gln)
barplot(pop.composition.gln, ylab="Proportion of population", legend=c("No GFP", "One copy GFP", "Two copy GFP", "Three+ copy GFP"),las=2, ylim=c(0,1), col=c("black", "grey", "light green", "dark green"), cex.names = 0.5,names.arg=sampleNames(flowGln))

```

#Step 5: Quality control

##Gates
```{r}
#Singlets gate
xyplot(FSC.A~FSC.H, data=flowData, xlim=c(0,3e6), ylim=c(0,3e6), filter=pg.singlets,  smooth=F, xbin=1024, stat=T, pos=0.5, abs=T, main = "First flowset - singlets gate")

###GLUCOSE###
#Debris gate
xyplot(SSC.A ~ FSC.A, data=flowGlu, displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=pg.nondebris.glu, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T,  main = "Glucose - nondebris gate")

#Non-fluorescent population gate
xyplot(FL1.A~FSC.A,data=filteredGluData, displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=pg.zcht, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T,  main = "Glucose - non GFP gate")

#ONe copy Fluorescent population gate
xyplot(FL1.A~FSC.A,data=filteredGluData, displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=pg.ocht, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T, main = "Glucose - One copy GFP gate")

#Two copy fluorescent population gate
xyplot(FL1.A~FSC.A,data=filteredGluData, displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=pg.twcht, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T, main = "Glucose - Two copy GFP gate")

#Three copy plus fluorescent population gate
xyplot(FL1.A~FSC.A,data=filteredGluData, displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=pg.thrcht, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T, main = "Glucose - Three+ copy GFP gate")

###UREA###
#Debris gate
xyplot(SSC.A ~ FSC.A, data=flowUr, displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=pg.nondebris.ur, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T,  main = "Urea - nondebris gate")

#Non-fluorescent population gate
xyplot(FL1.A~FSC.A,data=filteredUrData, displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=ur.zcht, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T,  main = "Urea - non GFP gate")

#ONe copy Fluorescent population gate
xyplot(FL1.A~FSC.A,data=filteredUrData, displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=ur.ocht, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T, main = "Urea - One copy GFP gate")

#Two copy fluorescent population gate
xyplot(FL1.A~FSC.A,data=filteredUrData, displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=ur.twcht, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T, main = "Urea - Two copy GFP gate")

#Three copy plus fluorescent population gate
xyplot(FL1.A~FSC.A,data=filteredUrData, displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=ur.thrcht, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T, main = "Urea - Three+ copy GFP gate")

###UREA###
#Debris gate
xyplot(SSC.A ~ FSC.A, data=flowGln, displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=pg.nondebris.gln, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T,  main = "Glutamine - nondebris gate")

#Non-fluorescent population gate
xyplot(FL1.A~FSC.A,data=filteredGlnData, displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=gln.zcht, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T,  main = "Glutamine - non GFP gate")

#ONe copy Fluorescent population gate
xyplot(FL1.A~FSC.A,data=filteredGlnData, displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=gln.ocht, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T, main = "Glutamine - One copy GFP gate")

#Two copy fluorescent population gate
xyplot(FL1.A~FSC.A,data=filteredGlnData, displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=gln.twcht, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T, main = "Glutamine - Two copy GFP gate")

#Three copy plus fluorescent population gate
xyplot(FL1.A~FSC.A,data=filteredGlnData, displayFilter=TRUE, xlim=c(0,5e6), ylim=c(0,6e5), filter=gln.thrcht, smooth=F, xbin=1024, stat=T, pos=0.5, abs=T, main = "Glutamine - Three+ copy GFP gate")
```


##Data transformation for visualization
```{r}
#In order to look at QC plots the data is transformed using the logicle transform, which is a log transform for high values that transitions to a linear transformation near zero values 

#This is simply for visualization purposes

lgcl <- logicleTransform(w = 0.5, t= 10000, m=4.5) #the parameters w,t, and m define the transformation parameters

#Tranformation applied to every channel except width and time
dataLGCLTransform <- transform(filteredGluData,'FSC.A' = lgcl(`FSC.A`), 'SSC.A' =lgcl(`SSC.A`), 'FL1.A' = lgcl(`FL1.A`), 'FL2.A' = lgcl(`FL2.A`), 'FL3.A' = lgcl(`FL3.A`), 'FL4.A' = lgcl(`FL4.A`),'FSC.H' = lgcl(`FSC.H`),'SSC.H' = lgcl(`SSC.H`),'FL1.H' = lgcl(`FL1.H`),'FL2.H' = lgcl(`FL2.H`),'FL3.H' = lgcl(`FL3.H`),'FL4.H' = lgcl(`FL4.H`)) 

dataLGCLTransform <- transform(filteredUrData,'FSC.A' = lgcl(`FSC.A`), 'SSC.A' =lgcl(`SSC.A`), 'FL1.A' = lgcl(`FL1.A`), 'FL2.A' = lgcl(`FL2.A`), 'FL3.A' = lgcl(`FL3.A`), 'FL4.A' = lgcl(`FL4.A`),'FSC.H' = lgcl(`FSC.H`),'SSC.H' = lgcl(`SSC.H`),'FL1.H' = lgcl(`FL1.H`),'FL2.H' = lgcl(`FL2.H`),'FL3.H' = lgcl(`FL3.H`),'FL4.H' = lgcl(`FL4.H`)) 

dataLGCLTransform <- transform(filteredGlnData,'FSC.A' = lgcl(`FSC.A`), 'SSC.A' =lgcl(`SSC.A`), 'FL1.A' = lgcl(`FL1.A`), 'FL2.A' = lgcl(`FL2.A`), 'FL3.A' = lgcl(`FL3.A`), 'FL4.A' = lgcl(`FL4.A`),'FSC.H' = lgcl(`FSC.H`),'SSC.H' = lgcl(`SSC.H`),'FL1.H' = lgcl(`FL1.H`),'FL2.H' = lgcl(`FL2.H`),'FL3.H' = lgcl(`FL3.H`),'FL4.H' = lgcl(`FL4.H`)) 


```

##Effect of time
```{r}
#The effect of time on signal (of which there shouldn't be any)
i <- 1
xyplot(FL1.A ~ Time, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(150,250), main = sampleNames(filteredGluData)[i])
i <- 2
xyplot(FL1.A ~ Time, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(150,250), main = sampleNames(filteredGluData)[i])
i <- 3
xyplot(FL1.A ~ Time, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(150,250), main = sampleNames(filteredGluData)[i])
```

##Plots of FL1 versus FSC
```{r}
i <- 1
xyplot(FL1.A ~ FSC.A, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(4,8), sampleNames(filteredGluData)[i])
i <- 2
xyplot(FL1.A ~ FSC.A, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(4,8), sampleNames(filteredGluData)[i])
i <- 3
xyplot(FL1.A ~ FSC.A, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(4,8), sampleNames(filteredGluData)[i])
```

##Plots of FSC versus SSC
```{r}
i <- 1
xyplot(SSC.A ~ FSC.A, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(4,8), ylim=c(4,8), sampleNames(filteredGluData)[i])
i <- 2
xyplot(SSC.A ~ FSC.A, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(4,8), ylim=c(4,8), sampleNames(filteredGluData)[i])
i <- 3
xyplot(SSC.A ~ FSC.A, data=dataLGCLTransform[i], smooth=F,  stat=T, pos=0.5, abs=T, xlim=c(4,8), ylim=c(4,8), sampleNames(filteredGluData)[i])
```

##Extract data for further processing
```{r}
summary.stats.glu <- as.data.frame(summary.stats.glu)
PopProportion_zeroGFPglu <- non.gfp.glu/non.debris.glu
PopProportion_oneGFPglu <- gfp.one.cells.glu/non.debris.glu
PopProportion_twoGFPglu <- gfp.two.cells.glu/non.debris.glu
PopProportion_threeGFPglu <- gfp.thr.cells.glu/non.debris.glu
alldataGLU <- cbind(summary.stats.glu, PopProportion_zeroGFPglu, PopProportion_oneGFPglu, PopProportion_twoGFPglu, PopProportion_threeGFPglu, non.debris.glu)
colnames(alldataGLU) <- c("FSC_median","FSC_mean", "FSC_sd","FL1_median", "FL1_mean","FL1_sd","normalizedGFP_median", "normalizedGFP_mean", "normalizedGFP_sd", "PopProp_0copy", "PopProp_1copy", "PopProp_2copy", "PopProp_3plus", "Cells counted")
###CHANGE COLUMN NAMES SO THEY ARE ALL THE SAME!!!!!!!

summary.stats.ur <- as.data.frame(summary.stats.ur)
PopProportion_zeroGFPur <- non.gfp.ur/non.debris.ur
PopProportion_oneGFPur <- gfp.one.cells.ur/non.debris.ur
PopProportion_twoGFPur <- gfp.two.cells.ur/non.debris.ur
PopProportion_threeGFPur <- gfp.thr.cells.ur/non.debris.ur
alldataur <- cbind(summary.stats.ur, PopProportion_zeroGFPur, PopProportion_oneGFPur, PopProportion_twoGFPur, PopProportion_threeGFPur, non.debris.ur)
colnames(alldataur) <- c("FSC_median","FSC_mean", "FSC_sd","FL1_median", "FL1_mean","FL1_sd","normalizedGFP_median", "normalizedGFP_mean", "normalizedGFP_sd", "PopProp_0copy", "PopProp_1copy", "PopProp_2copy", "PopProp_3plus", "Cells counted")

summary.stats.gln <- as.data.frame(summary.stats.gln)
PopProportion_zeroGFPgln <- non.gfp.gln/non.debris.gln
PopProportion_oneGFPgln <- gfp.one.cells.gln/non.debris.gln
PopProportion_twoGFPgln <- gfp.two.cells.gln/non.debris.gln
PopProportion_threeGFPgln <- gfp.thr.cells.gln/non.debris.gln
alldatagln <- cbind(summary.stats.gln, PopProportion_zeroGFPgln, PopProportion_oneGFPgln, PopProportion_twoGFPgln, PopProportion_threeGFPgln, non.debris.gln)
colnames(alldatagln) <- c("FSC_median","FSC_mean", "FSC_sd","FL1_median", "FL1_mean","FL1_sd","normalizedGFP_median", "normalizedGFP_mean", "normalizedGFP_sd", "PopProp_0copy", "PopProp_1copy", "PopProp_2copy", "PopProp_3plus", "Cells counted")

ALLDATA <- rbind(alldataGLU, alldataur, alldatagln)

#Save the current dataframe
write.table(ALLDATA, file="DataOut_05252017_MIDTPT.csv", row.names=TRUE, quote=F, sep="\t")

```




